---
title: "Oak Whittle"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**This code is to create the data sheets needed to collapse the oaks recorded as more than 1 tree into a single tree. That work will have to be done manually in excel**

# Packages
```{r}
library(tidyverse)
library(here)
library(janitor)
```

# Load Data
```{r}
# FIA 
tree_data   <- read.csv(here("./FIA_Data/Online_FIA_Data/2001_2018_Raw_FIA_Data/CA_TREE.csv"))
fia         <- read.csv(here("./FIA_Data/Final FIA data and code used for analysis/final_binded_fia_ypmc_data.csv")) %>% 
  select(unique_plot) %>% 
  distinct(.keep_all = T) # use this to bind to the oaks prelim to get just oaks in the plots we will analyze for the analysis


# USFS FSVeg
anf <- read.csv(here("./FIA_Data/USFS_intensified_and_FIA_data/tree_data/ANF_FSVEG_TXPT_V.csv"))
cnf <- read.csv(here("./FIA_Data/USFS_intensified_and_FIA_data/tree_data/cNF_FSVEG_TXPT_V.csv"))
lpnf <- read.csv(here("./FIA_Data/USFS_intensified_and_FIA_data/tree_data/LPF_FSVEG_TXPT_V.csv"))
sbnf <- read.csv(here("./FIA_Data/USFS_intensified_and_FIA_data/tree_data/BDF_FSVEG_TXPT_V.csv"))

fsveg_prelim <- rbind(anf, cnf, lpnf, sbnf) %>% clean_names()

```

**FIA METADATA**
most recent *metadata* version:
https://www.fia.fs.fed.us/library/database-documentation/current/ver80/FIADB%20User%20Guide%20P2_8-0.pdf

Data frame desciptions (digital pg 34) from *older metadata* https://www.fia.fs.fed.us/library/database-documentation/current/ver60/FIADB%20User%20Guide%20P2_6-0-2_final-opt.pdf


**USFS INTENSIFIED PLOT METADATA**
https://www.fs.fed.us/r5/rsl/projects/inventory/fsveg-tree.shtml


# FIA Oaks

## Get and Export Oaks

Make separate data sheet to collapse oaks recorded as multiple trees into a single tree.

This is initial work done to create the csv added above in the load data section
```{r}
# filter/tidy
oaks_fia <- tree_data %>%
  unite(unique_plot, c("COUNTYCD", "PLOT", "INVYR"), sep = "-", remove=T) %>%
  clean_names() %>%
  filter(dia >= 4) %>%
  filter(spcd %in% c(805, 818)) %>% # 805 = QUCH, 818 = QUKE
  mutate(species = case_when(
    spcd == 805 ~ "QUCH",
    spcd == 818 ~ "QUKE")) %>%
  filter(statuscd == 1) %>% # keep only alive trees/remove dead trees
  dplyr::select(unique_plot, subp, tree, azimuth, dist, dia, species)

# join to get oaks just in the fia plots used in our analysis
oaks_join <- inner_join(fia, oaks_fia, by = "unique_plot")

# export to manually group oaks
#write.csv(oaks_join, here("./FIA Data/Oak Analysis/fia_analysis oaks join.csv"), row.names = F)

```
dist units = feet


# FSVEG Data

```{r}
oaks_fsveg <- fsveg_prelim %>% 
  filter(dbhi >= 4) %>% 
  filter(spec %in% c("QUCH2", "QUKE")) %>% 
  filter(lvdd == "L") %>% # keep alive trees/remove dead trees
  select(plot, tree, tcnt, spec, dbhi)
```

columns that coule be helpful

- tcnt: treecount --> only has values of 1 :(
