---
title: "vtm_shrubs"
author: "Leana Goetze"
date: "1/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

# attach packages
```{r}
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)

```

# Read in final list of YPMC plots 
```{r}
vtm_plots <- read.table(here::here("vtm_data", "vtm_dataset_750m_buffer_FINAL.txt"))%>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0))

# check that this is correct
ypmc_plots <-unique(vtm_plots$plotkey)
# 195-yes!
```

# Read in raw brush datasheet
```{r}
# load (manipulated) original datasets. Manipulations: The plotkey columns have been deleted from the "brush", "flatbrush", "tree", and "flattrees" datasets and the plotkey values have been placed in quotes on the "plot" datasheet (see notes.docx for more details). --> these manipulated datasheets were accidently deleted. To rerun this code will have to recreate them. 

raw_brush<- read_delim(here("vtm_data", "vtm-plotdata-brush_2.csv"), delim = ",") %>% 
  janitor::clean_names()
raw_plot<- read_delim(here("vtm_data","vtm-plotdata-plot_2.csv"), delim = ",") %>% 
  janitor::clean_names() 

# remove quotation marks around the values in the plotkey column on the plot datasheet
original_plot<- raw_plot %>% 
  mutate(new_plotkey = (str_remove(plotkey, pattern = '"'))) %>% 
  select(-plotkey) %>% 
  mutate(plotkey = (str_remove(new_plotkey, pattern = '"'))) %>% 
  select(-new_plotkey)

# reorder columns
original_plot <-original_plot [,c(40, 1:39)]

# create df representing "pid" and corresponding "plotkeys" values
pid_plotkey <- original_plot %>% 
  select(pid, plotkey)

# match pid column with corresponding plotkeys values in original datasets
# reorder columns
original_brush<- raw_brush %>% 
  left_join(pid_plotkey, by = "pid")
original_brush<-original_brush[,c(20, 1:19)]
```

# Filter brush dataframe for YPMC plots of interest
```{r}
ypmc_brush <- original_brush %>% 
  filter (plotkey %in% ypmc_plots)

unique(ypmc_brush$plotkey)
# we have brush data on 182 plots out of 195 plots
```

# filter brush dataframe for only shrub species
```{r}
# filter out anything but shrubs
# NOTE: quercus wislizenii has been included wih shrubs (can be either a shrub or tree)
filtered_ypmc_brush <- ypmc_brush %>% filter(!code %in% c("Lit", "X", "Gr", "Lit, Conifer", "Lit, Cp", "Lit, S, W1", "Lit, Con", "Lit, BS", "Lit, Cl", "Lit, Jo", "Lit, S, C", "Lit, W1", "Lit, Popu", "Lit, I", "Lit, SP, JP", "Lit, Sp, Yp", "Lit, Y, W1", "W1", "Acg", "I", "Fern", "BS", "Pta", "Lit, C", "Lit, B", "Lit, W", "Lit, Oak")) %>% 
  filter(!genus %in% c("pinus", "grass", "barren")) %>% 
  filter(!lifeform %in% c("rock")) %>% 
  filter(!name %in% c("Unmatched species code"))

# convert to lowercase species names
filtered_ypmc_brush$name = tolower(filtered_ypmc_brush$name)

# attach common names and further filter out species that are not shrubs
final_ypmc_brush <-filtered_ypmc_brush %>% 
  mutate(common_name = case_when(name == "adenostoma fasciculatum " ~ "chamise, greasewood,huutah",
                                 name == "amelanchier alnifolia " ~ "saskatoon, pacific serviceberry, western serviceberry, alder-leaf shadbush, dwarf shadbush, chuckley pear, western juneberry",
                                 name == "amorpha californica " ~ "california false indigo",
                                 name == "arctostaphylos canescens " ~ "hoary manzanita",
                                 name == "arctostaphylos drupacea " ~ "pinkbracted manzanita",
                                 name == "arctostaphylos glandulosa " ~ "eastwood manzanita",
                                 name == "arctostaphylos glauca " ~ "bigberry manzanita",
                                 name == "arctostaphylos parryana pinetorum" ~ "parry manzanita",
                                 name == "arctostaphylos parryana " ~ "parry manzanita",
                                 name == "arctostaphylos pungens " ~ "pointleaf manzanita",
                                 name == "arctostaphylos tomentosa " ~ "woolly leaf manzanita",
                                 name == "artemisia tridentata " ~ "big sagebrush",
                                 name == "artemisia tridentata parishii or apocynum androsaemifolium pumilum" ~ "big sagebrush or spreading dogbane",
                                 name == "balsamorhiza sagittata " ~ "remove",
                                 name == "castanopsis sempervirens " ~ "bush chinquapin",
                                 name == "ceanothus cordulatus " ~ "mountain whitethorn",
                                 name == "ceanothus divaricatus " ~ "hairy ceanothus",
                                 name == "ceanothus greggii " ~ "desert ceanothus",
                                 name == "ceanothus integerrimus " ~ "deer brush",
                                 name == "ceanothus prostratus " ~ "prostrate ceanothus",
                                 name == "cercocarpus betuloides " ~ "mountain mahogany",
                                 name == "cercocarpus ledifolius " ~ "curl-leaf mountain mahogany",
                                 name == "chrysothamnus nauseosus " ~ "Chamisa, rubber rabbitbrush,gray rabbitbrush",
                                 name == "clematis pauciflora " ~ "remove",
                                 name == "encelia actoni " ~ "acton brittlebush, acton encelia",
                                 name == "eriogonum fasciculatum " ~ "california buckwheat",
                                 name == "eriogonum kennedyi " ~ "remove",
                                 name == "eriogonum sp. " ~ "wild buckwheat?",
                                 name == "fremontia californica " ~ "california flannelbush",
                                 name == "garrya veatchii " ~ "canyon silktasse, Veatch silktassel",
                                 name == "holodiscus discolor " ~ "oceanspray, creambush, ironwood",
                                 name == "leptodactylon pungens " ~ "granite prickly-phlox, granite gilia",
                                 name == "lonicera conjugialis " ~ "purpleflower honeysuckle",
                                 name == "lupinus sp. " ~ "remove",
                                 name == "prunus emarginata " ~ "bitter cherry, oregon cherry",
                                 name == "prunus ilicifolia " ~ "hollyleaf cherry, evergreen cherry",
                                 name == "rhamnus californica " ~ "california coffeeberry, california buckthorn",
                                 name == "rhus trilobata " ~ "skunkbush sumac, sourberry, skunkbush, Squawbush, three-leaf sumac",
                                 name == "ribes cereum " ~ "wax currant",
                                 name == "ribes nevadense " ~ "sierra currant, mountain pink currant",
                                 name == "ribes roezlii " ~ "sierra gooseberry",
                                 name == "ribes sp. " ~ "currants, gooseberries",
                                 name == "salix sp. " ~ "remove",
                                 name == "salvia carnosa " ~ "purple sage",
                                 name == "salvia carnosa compacta" ~ "purple sage??",
                                 name == "salvia palmeri " ~ "remove",
                                 name == "symphoricarpos mollis " ~ "creeping snowberry, southern california snowberry,trip vine",
                                 name == "quercus wislizenii " ~ "Interior live oak",
                                 name == "quercus dumosa " ~ "coastal sage scrub oak",
                                 name == "quercus chrysolepis " ~ "canyon live oak",
                                 name == "quercus kelloggii  " ~ "california black oak",
                                 name == "quercus agrifolia  " ~ "coast live oak",
                                 TRUE ~ "none")) %>% 
  filter(!common_name == "remove")

# rename Castanopsis sempervirens to Chrysolepis sempervirens 
final_ypmc_brush$genus[final_ypmc_brush$genus == "castanopsis"] <-"chrysolepis"
final_ypmc_brush$name[final_ypmc_brush$name == "castanopsis sempervirens "] <-"chrysolepis sempervirens"


# check stats on final shrub list in YPMC plots
unique(final_ypmc_brush$plotkey) 
unique(final_ypmc_brush$code) 
# 138/195 plots
# 47 shrub species

# create list of shrubs
list <- final_ypmc_brush %>% 
  select(name, common_name) 
# make kable table
  list %>% 
    kable(col.names = c("Scientific Name", 
                      "Common Name")) %>% 
  kable_styling(bootstrap_options = "bordered", 
                full_width = F,
                position = "left") 
  # save_kable("shrubs.png")
```

# filter final shrubs dataframe to make more manageable
```{r}
final_shrubs <- final_ypmc_brush %>% 
  select(plotkey, code, percent, genus, species, name, common_name) %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0))
# NOTE: A percent of "0" means the percent is unknown. It was marked as "NA"
```

# attach plot (and additional) data
```{r}
# select columns of interest from plot data and filter for plotkeys of interest (plots in YPMC buffer)
plot_subset<-vtm_plots %>% 
  select(plotkey, national_forest, national_forest_name, private_land, state_park, wilderness, range, elevation, exposure, slope_percent, date, notes, latitude, longitude) %>% 
   filter(plotkey %in% final_shrubs$plotkey)

# delete duplicate rows
plot_subset_final <-unique(plot_subset)

unique(plot_subset_final$plotkey)
#138-yes!

combined_shrub_data <-final_shrubs %>% 
  full_join(plot_subset_final, by = "plotkey")

sum(combined_shrub_data$percent)
# 2717 total subplots recorded (not including NA)
```

# total species dominance
```{r}
total_species_dominance <-combined_shrub_data %>% 
  group_by(common_name, name) %>% 
  summarise(total_count = sum(percent)) 

#check class
class(total_species_dominance$total_count)
# numeric

# sort by descending order
sorted <-total_species_dominance[order(-total_species_dominance$total_count),]

# create percent column
shrub_dataframe <-sorted %>% 
  mutate(percent = ((total_count/(sum(total_species_dominance$total_count))))*100)

# sum(total_species_dominance$total_count)
#2717 total shrub counts

final_shrub_dataframe <-shrub_dataframe

# NOTE: the "percent" column is the percent that the given species dominates the shrub observations, not the percent at which it covers the landscape. For instance: ceanothus cordulatus made up 371 of the 2717 subplots dominated by shrubs, or 13.7%. 
```

# Overall top ten dominant shrubs
```{r}
# top ten dominant shrubs
top_ten_overall <-final_shrub_dataframe %>% 
  head(10) %>% 
  mutate(percent = round(percent, 2)) %>% 
  select("common_name", "name", "percent", "total_count" )

# make table
top_ten_overall %>% 
    kable(col.names = c("Common Name", 
                      "Scientific Name",
                      "Percent Dominance",
                      "Sample Size")) %>% 
  kable_styling(bootstrap_options = "bordered", 
                full_width = F,
                position = "left") %>% 
  row_spec(2, background = "#CCD5E1") %>% 
  row_spec(4, background = "#CCD5E1") %>%
  row_spec(6, background = "#CCD5E1") %>%
  row_spec(8, background = "#CCD5E1") %>% 
  row_spec(10, background = "#CCD5E1") %>% 
  row_spec(1, background = "#EAF1FB") %>% 
  row_spec(3, background = "#EAF1FB") %>% 
  row_spec(5, background = "#EAF1FB") %>% 
  row_spec(7, background = "#EAF1FB") %>% 
  row_spec(9, background = "#EAF1FB")
  # save_kable("top_ten_species.png")
```

# top three at each elevation gradient
```{r}
# create elevation classes
elevation_dataset <-combined_shrub_data %>%
  mutate(elevation_m = (elevation/3.28084))%>% #convert from ft. to m
  mutate(elevation_class = case_when (elevation_m <= 499 ~ "<499",
                                    elevation_m > 500 & elevation_m <999.99 ~ "500–999", 
                                    elevation_m > 1000 & elevation_m <1499.99 ~ "1000–1499",
                                    elevation_m > 1500 & elevation_m <1999.99 ~ "1500–1999",
                                    elevation_m > 2000 & elevation_m <2499.99 ~ "2000–2499",
                                    elevation_m >= "2500" ~ ">2500",
                                    TRUE ~ "none")) %>% 
  filter(!plotkey == "163BB28") %>% # no elevation data for this plotkey  
  filter(!national_forest_name == "N/A") %>% 
  group_by(elevation_class, common_name, name) %>% 
  summarise(total_count = sum(percent))

# sum total count for each elevation class
elevation_dataset_2 <- elevation_dataset %>% 
  group_by(elevation_class) %>% 
  mutate(sum = sum(total_count))

# create percent shrub dominance for each elevation class
elevation_dataset_3 <- elevation_dataset_2 %>% 
  group_by(elevation_class, common_name) %>% 
  mutate(percent = (total_count/sum)*100)

# order by descending percent values and select the top greatest values for each elevation class
elevation_dataset_4 <- elevation_dataset_3 %>% 
  group_by(elevation_class) %>% 
  slice_max(order_by = percent, n = 3)

top_five_elevation_class <- elevation_dataset_4 %>% 
  select(elevation_class, common_name, name, percent, total_count)%>% 
  mutate(percent = round(percent, 2))

# fix order of elevation bands
level.order <- c("1000–1499","1500–1999", "2000–2499", ">2500")

top_three_elevation_class <-top_five_elevation_class[order(factor(top_five_elevation_class$elevation_class, levels = level.order)),]

# make table
top_three_elevation_class %>% 
    kable(col.names = c("Elevation (m)", 
                      "Common Name",
                      "Scientific Name",
                      "Percent Dominance",
                      "Sample Size")) %>% 
  kable_styling(bootstrap_options = "bordered", 
                full_width = F,
                position = "left") %>% 
  row_spec(1, background = "#DBE2E6") %>% 
  row_spec(2, background = "#DBE2E6") %>%
  row_spec(3, background = "#DBE2E6") %>%
  row_spec(4, background = "#BCC2C6") %>% 
  row_spec(5, background = "#BCC2C6") %>%
  row_spec(6, background = "#BCC2C6") %>%
  row_spec(7, background = "#DBE2E6") %>%
  row_spec(8, background = "#DBE2E6") %>%
  row_spec(9, background = "#DBE2E6") %>% 
  row_spec(10, background = "#BCC2C6") %>%
  row_spec(11, background = "#BCC2C6") %>%
  row_spec(12, background = "#BCC2C6") 
  # save_kable("top_three_elevation.png")
```

# top three from each NF
```{r}
nf_dataset <-combined_shrub_data%>% 
  group_by(national_forest_name, common_name, name) %>% 
  filter(!national_forest_name == "N/A") %>% 
  summarise(total_count = sum(percent)) 

# sum total count for each elevation class
nf_dataset_2 <- nf_dataset %>% 
  group_by(national_forest_name) %>% 
  mutate(sum = sum(total_count))

# create percent shrub dominance for each elevation class
nf_dataset_3 <- nf_dataset_2 %>% 
  group_by(national_forest_name, common_name) %>% 
  mutate(percent = (total_count/sum)*100)

# order by descending percent values and select the top greatest values for each elevation class
nf_dataset_4 <- nf_dataset_3 %>% 
  group_by(national_forest_name) %>% 
  slice_max(order_by = percent, n = 3)

top_three_nf <- nf_dataset_4 %>% 
  select(national_forest_name, common_name, name, percent, total_count)%>% 
  mutate(percent = round(percent, 2))

# make table
top_three_nf %>% 
    kable(col.names = c("National Forest", 
                      "Common Name",
                      "Scientific Name",
                      "Percent Dominance", 
                      "Sample Size")) %>% 
  kable_styling(bootstrap_options = "bordered", 
                full_width = F,
                position = "left") %>% 
  row_spec(1, background = "#DBE2E6") %>% 
  row_spec(2, background = "#DBE2E6") %>%
  row_spec(3, background = "#DBE2E6") %>%
  row_spec(4, background = "#BCC2C6") %>%
  row_spec(5, background = "#BCC2C6") %>% 
  row_spec(6, background = "#BCC2C6") %>%
  row_spec(7, background = "#DBE2E6") %>%
  row_spec(8, background = "#DBE2E6") %>% 
  row_spec(9, background = "#DBE2E6") %>% 
  row_spec(10, background = "#BCC2C6") %>%
  row_spec(11, background = "#BCC2C6") %>% 
  row_spec(12, background = "#BCC2C6")
  # save_kable("top_three_nf.png")
```

# top ten frequenct genuses
```{r}
# summarise genuses
genuses_dataset<- combined_shrub_data %>% 
  group_by(genus) %>% 
  summarise(total_count = sum(percent)) 

genuses_dataset_2 <- genuses_dataset %>% 
  mutate(percent = ((total_count/(sum(genuses_dataset$total_count))))*100)

# reorder column
final_genuses_dataset<-genuses_dataset_2[order(-genuses_dataset_2$percent),]

# top ten frequent genuses
top_ten_genuses <-final_genuses_dataset %>% 
  head(10) %>% 
  mutate(percent = round(percent, 2)) %>% 
  select(genus, percent, total_count)

# make table
top_ten_genuses  %>% 
    kable(col.names = c("Genus",
                      "Percent Dominance", 
                      "Sample Size")) %>% 
  kable_styling(bootstrap_options = "bordered", 
                full_width = F,
                position = "left") %>% 
  row_spec(2, background = "#CCD5E1") %>% 
  row_spec(4, background = "#CCD5E1") %>%
  row_spec(6, background = "#CCD5E1") %>%
  row_spec(8, background = "#CCD5E1") %>% 
  row_spec(10, background = "#CCD5E1") %>% 
  row_spec(1, background = "#EAF1FB") %>% 
  row_spec(3, background = "#EAF1FB") %>% 
  row_spec(5, background = "#EAF1FB") %>% 
  row_spec(7, background = "#EAF1FB") %>% 
  row_spec(9, background = "#EAF1FB")  
  # save_kable("top_ten_genuses.png")


# top ten frequent genuses- remove NA
top_ten_genuses_no_na <-final_genuses_dataset %>% 
  filter(!genus == "NA") %>% 
  head(10) %>% 
  mutate(percent = round(percent, 2)) %>% 
  select(genus, percent, total_count)

# make table
top_ten_genuses_no_na  %>% 
    kable(col.names = c("Genus",
                      "Percent Dominance", 
                      "Sample Size")) %>% 
  kable_styling(bootstrap_options = "bordered", 
                full_width = F,
                position = "left") %>% 
  row_spec(2, background = "#CCD5E1") %>% 
  row_spec(4, background = "#CCD5E1") %>%
  row_spec(6, background = "#CCD5E1") %>%
  row_spec(8, background = "#CCD5E1") %>% 
  row_spec(10, background = "#CCD5E1") %>% 
  row_spec(1, background = "#EAF1FB") %>% 
  row_spec(3, background = "#EAF1FB") %>% 
  row_spec(5, background = "#EAF1FB") %>% 
  row_spec(7, background = "#EAF1FB") %>% 
  row_spec(9, background = "#EAF1FB")
  # save_kable("top_ten_genuses_no_na.png")

```

# by elevation and national forest produced way too small of sample sizes--> do genus insead
# by genus and elevation
```{r}
elevation_genus <-combined_shrub_data %>%
  mutate(elevation_m = (elevation/3.28084))%>% #convert from ft. to m
  mutate(elevation_class = case_when (elevation_m <= 499 ~ "<499",
                                    elevation_m > 500 & elevation_m <999.99 ~ "500–999", 
                                    elevation_m > 1000 & elevation_m <1499.99 ~ "1000–1499",
                                    elevation_m > 1500 & elevation_m <1999.99 ~ "1500–1999",
                                    elevation_m > 2000 & elevation_m <2499.99 ~ "2000–2499",
                                    elevation_m >= "2500" ~ ">2500",
                                    TRUE ~ "none"))%>% 
  filter(!plotkey == "163BB28") %>%  # no elevation data for this plotkey  %>% 
  filter(!national_forest_name == "N/A")%>% 
  group_by(elevation_class, genus)%>% 
  summarise(total_count = sum(percent))

# sum total count for each elevation class --> this is not needed
elevation_genus_2 <- elevation_genus %>% 
  group_by(elevation_class) %>% 
  mutate(sum = sum(total_count))

#create percent shrub dominance for each elevation class
elevation_genus_3 <- elevation_genus_2 %>% 
  group_by(elevation_class, genus) %>% 
  mutate(percent = (total_count/sum)*100)

#order by descending percent values and select the top greatest values for each elevation class
elevation_genus_4 <- elevation_genus_3 %>% 
  group_by(elevation_class) %>% 
  slice_max(order_by = percent, n = 3)%>% 
  select(elevation_class, genus, percent, total_count ) %>%
  mutate(percent = round(percent, 2)) 

# fix order of elevation bands
level.order <- c("1000–1499","1500–1999", "2000–2499", ">2500")

elevation_genus_final<-elevation_genus_4[order(factor(elevation_genus_4$elevation_class, levels = level.order)),]

# make table
top_three_genuses_by_elevation <- elevation_genus_final %>% 
    kable(col.names = c("Elevation (m)",
                        "Genus",
                      "Percent Dominance",
                      "Sample Size")) %>% 
  kable_styling(bootstrap_options = "bordered", 
                full_width = F,
                position = "left") %>% 
  row_spec(1, background = "#CCD5E1") %>% 
  row_spec(2, background = "#CCD5E1") %>%
  row_spec(3, background = "#CCD5E1") %>%
  row_spec(4, background = "#EAF1FB") %>% 
  row_spec(5, background = "#EAF1FB") %>% 
  row_spec(6, background = "#EAF1FB") %>% 
  row_spec(7, background = "#CCD5E1") %>% 
  row_spec(8, background = "#CCD5E1") %>% 
  row_spec(9, background = "#CCD5E1") %>% 
  row_spec(10, background = "#EAF1FB") %>% 
  row_spec(11, background = "#EAF1FB") %>% 
  row_spec(12, background = "#EAF1FB")
  # save_kable("top_three_genuses_by_elevation.png")

top_three_genuses_by_elevation
```

