---
title: "Untitled"
author: "Leana Goetze"
date: "11/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Upload packages
library(tidyverse)
library(tidyr)
library(here)
```

### I. VTM
```{r}
# read in data
vtm_data <- read.table(here::here("vtm_data", "vtm_dataset_750m_buffer_FINAL.txt"))%>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))%>%
  mutate(n= dbh_4_11 + dbh_12_23 + dbh_24_35 +dbh_36)
```

### Canopy Cover Overview
## Equations

# 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
# 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
# 4.4.1.4: CW = HT * s1 (HT <4.5' and DBH<DBHT)
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT)

# CW is maximum tree crown width
# DBH is tree diameter at breast height
# DBH(subT) is threshold diameter 

## Table (from...) Coefficients and equation reference for crown width equations {4.4.2.1}-{4.4.2.5} in the WS variant

# Sugar Pine (PILA), Equation 4.4.2.1, DBH(subT) = 7.4 / SP
# White fir (ABCO), Equation 4.4.2.1, DBH(subT) = 5 / WF
# Incense cedar (CADE), Equation 4.4.2.1, DBH(subT) = 5 / IC
# Jeffrey Pine (PIJE), Equation 4.4.2.2, DBH(subT) = 5 / JP
# Ponderosa Pine (PIPO), Equation 4.4.2.2, DBH(subT) = 5 / PP
# Coulter Pine (PICO), Equation 4.4.2.2, DBH(subT) = 5 / CP
# Bigcone Spruce (PSMA), Equation 4.4.2.1, DBH (subT) = 5 / BD
# Canyon Live Oak (QUCH), Equation 4.4.2.1, DBH(subT) = 5 / CY
# Black Oak (QUKE), Equation 4.4.2.1, DBH(subT) = 5 / BO

```{r}
## Constants for equations 
# --------------------------------------------------
# Sugar Pine (PILA) constants:
pila_d1 <- 3.5
pila_d2 <- 0.338
pila_a1 <- -1.476
pila_a2 <- 1.01
pila_s1 <- 0.7778

# White fir (ABCO) constants:
abco_d1 <- 3.26
abco_d2 <- 1.103
abco_a1 <- 5.82
abco_a2 <- 0.591
abco_s1 <- 0.7778

# Incense Cedar (CADE) constants:
cade_d1 <- 3.5
cade_d2 <- 1.192
cade_a1 <- 7.11
cade_a2 <- 0.47
cade_s1 <- 0.7778

# Jeffrey Pine (PIJE) constants: 
pije_d1 <- 3.5
pije_d2 <- 0.5754
pije_a1 <- 1.52
pije_a2 <- 0.891
pije_s1 <- 0.7778

# Ponderosa Pine (PIPO) constants:
pipo_d1 <- 3.77
pipo_d2 <- 0.7756
pipo_a1 <- 2.24
pipo_a2 <- 0.763
pipo_s1 <- 0.7778

# Coulter Pine (PICO) constants:
pico_d1 <- 3.5
pico_d2 <- 1.7618
pico_a1 <- 3.9347
pico_a2 <- 0.7086
pico_s1 <- 0.7778

# Bigcone Spruce (PSMA) constants
psma_d1 <- 3.5
psma_d2 <- 5.5672
psma_a1 <- 27.030
psma_a2 <- 0.8612
psma_s1 <- 0.7778

# Canyon Live Oak (QUCH) constants:
quch_d1 <- 2.5
quch_d2 <- 2.19
quch_a1 <- 5
quch_a2 <- 1.69
quch_s1 <- 0.5556

# Black Oak (QUKE) constants:
quke_d1 <- 2.5
quke_d2 <- 2.7
quke_a1 <- 10
quke_a2 <- 1.2
quke_s1 <- 0.5556

pi <- 3.1415926535 # Leana added this constant
e <- 2.718281828 # Leana added this constant
```


```{r}
## Anne-Marie's midpoints for 36+ dbh
# vtm_basal_largest <- vtm_tidy %>% 
#   filter(dbh_class == "dbh_36") %>% 
#   mutate(dbh_final = case_when(
#     species == "ABCO" ~ 106 + 14,-->47.2
#     species == "CADE" ~ 162,-->63.8
#     species == "PICO" ~ 99 + 5,-->40.9
#     #species == "PIJE" ~ 104 + 12,-->45.7
#     species == "PILA" ~ 108 + 15,-->48.4
#     #species == "PIPO" ~ 110 + 18,-->50.4
#     species == "Yellow Pine" ~ 104 + 13,-->46.1
#     species == "PSMA" ~ 107 + 13,-->47.2
#     species == "QUCH" ~ 98 - 6,-->36.2
#     species == "QUKE" ~ 93-->36.6
```

```{r}
# How to go about canopy cover calculation
# From Hugh NRV: "First, we used species-specific equations for crown-width (above) to estimate crown area, using the midpoint of thee size class in our calculations, then multiplying by the number of trees in that size class. Percentage of cover was obtained by summing across the size classes and standardizing to square meters, then dividing the result by 10,000 (number of square meters in a hectare) and multiplying by 100 for percentage. 
# From Keyser 2010: "Crown width for each tree is reported in the tree list ouput table and used for percent canopy cover (PCC) calculations. Crown width is calculated by using equations [4.4.2.1] - [4.4.2.5]. If a tree has a DBH greater than or equal to its threshold diameter (given as DBH(subT), then it uses equations [4.4.2.1], [4.4.2.2], or [4.4.2.3] depending on the species. If a tree has a DBH less than its threshold diameter, then it uses equation [4.4.2.4] or [4.4.2.5], depending on the height of the tree. )
```


# PIPO Canopy Cover 
```{r}
# Use Equation 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
# DBH(subT) is 5 and our lowest dbh is 7.5 since we are taking the midpoint of each class. 

# Select only PIPO & select only needed columns
# Pivot_longer to get size classes as observations
# get rid of zeros (goes from 328 observations to 198). When uncount, goes from 198 to 865.

vtm_pipo <- vtm_data %>%
  filter(abbreviation == "PIPO") %>%
  select(plotkey, abbreviation, n,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent)%>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "46.1")) # used Anne-Marie's measurements. Used the yellow pine value for pipo and pije, but we should double check this!

# change dbh_class & canopy_cover to numeric
vtm_pipo$dbh_class <- as.numeric(vtm_pipo$dbh_class)
# str(vtm_pipo)
vtm_pipo <- mutate(vtm_pipo, canopy_width = (pipo_a1 * (dbh_class^(pipo_a2)))) # Leana renamed canopy_cover to canopy_width because this is what the equation is calculating
vtm_pipo$canopy_width <- as.numeric(vtm_pipo$canopy_width)
# note: canopy_width is in feet (https://sourceforge.net/p/open-fvs/wiki/FVS_API/?version=2)

# Questions: 
# 1. What do we want to do with the midpoint for 36+ class? (hgw) I forgot, did we discuss this with Bruce/Anne-Marie at the meeting? I'm guessing it would be hard to find literature on the largest recorded dbh per species from the 1930s...perhaps we could use the largest dbh per species from the FIA data and use that as the bound? I know it's not the best, but it is some measure of how big each species can get. 

## ------------------------------------------------------------------------------------------------
# Leana added this section. It is best to calculate canopy area here, and not after grouping the data, because canopy area should be calculated based off of the canopy_width and not the sum of all canopy_widths. Ex. The canopy area of a tree with a canopy width of 10 is roughly 78.5. Let's say we have 5 trees in this size class, then the total canopy area of all five trees would be (78.5 * 5), which is roughly 396. This is not equivalent to adding the canopy widths (10*5), and then finding the area based off of this width. The area of a tree with crown width of 50 is roughly 1963. I hope that makes sense!

# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_pipo_canopy <- vtm_pipo %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) # Hannah, this was correct. I just deleted the part where you converted to feet, because I believe crown width is already in feet (this is what we discussed with Bruce). 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_pipo_sum <- vtm_pipo_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) # checked a couple of plots against the vtm_pipo dataframe and they added up to the correct amount.  
```

# PIJE Canopy Cover 
```{r}
# Equation for PIJE: # 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
vtm_pije <- vtm_data %>%
  filter(abbreviation == "PIJE") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "46.1")) # used Anne-Marie's measurements. Used the yellow pine value for pipo and pije, but we should double check this!

vtm_pije$dbh_class <- as.numeric(vtm_pije$dbh_class)
vtm_pije <- mutate(vtm_pije, canopy_width = (pije_a1 * (dbh_class^(pije_a2))))
vtm_pije$canopy_width <- as.numeric(vtm_pije$canopy_width)

## ------------------------------------------------------------------------------------------------
# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_pije_canopy <- vtm_pije %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_pije_sum <- vtm_pije_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# PILA Canopy Cover
```{r}
# Use Equation: # 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
vtm_pila <- vtm_data %>%
  filter(abbreviation == "PILA") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "48.4"))

vtm_pila$dbh_class <- as.numeric(vtm_pila$dbh_class)
vtm_pila <- mutate(vtm_pila, canopy_width = (pila_a1 + pila_a2 * (dbh_class)))
vtm_pila$canopy_width <- as.numeric(vtm_pila$canopy_width)

## ------------------------------------------------------------------------------------------------# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_pila_canopy <- vtm_pila %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_pila_sum <- vtm_pila_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# PICO Canopy Cover
```{r}
# Use Equation 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
vtm_pico <- vtm_data %>%
  filter(abbreviation == "PICO3") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "40.9"))

vtm_pico$dbh_class <- as.numeric(vtm_pico$dbh_class)
vtm_pico <- mutate(vtm_pico, canopy_width = (pico_a1 * (dbh_class^(pico_a2))))
vtm_pico$canopy_width <- as.numeric(vtm_pico$canopy_width)

## ------------------------------------------------------------------------------------------------# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_pico_canopy <- vtm_pico %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_pico_sum <- vtm_pico_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# ABCO Canopy Cover
```{r}
# # White fir (ABCO), Equation 4.4.2.1, DBH(subT) = 5
# # 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)

vtm_abco <- vtm_data %>%
  filter(abbreviation == "ABCO") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "47.2"))

vtm_abco$dbh_class <- as.numeric(vtm_abco$dbh_class)
vtm_abco <- mutate(vtm_abco, canopy_width = (abco_a1 + abco_a2 * (dbh_class)))
vtm_abco$canopy_width <- as.numeric(vtm_abco$canopy_width)

## ------------------------------------------------------------------------------------------------# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_abco_canopy <- vtm_abco %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_abco_sum <- vtm_abco_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# CADE Canopy Cover
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)

vtm_cade <- vtm_data %>%
  filter(abbreviation == "CADE") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "63.8"))

vtm_cade$dbh_class <- as.numeric(vtm_cade$dbh_class)
vtm_cade <- mutate(vtm_cade, canopy_width = (cade_a1 + cade_a2 * (dbh_class)))
vtm_cade$canopy_width <- as.numeric(vtm_cade$canopy_width)

## ------------------------------------------------------------------------------------------------# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_cade_canopy <- vtm_cade %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_cade_sum <- vtm_cade_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# PSMA canopy cover
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
vtm_psma <- vtm_data %>%
  filter(abbreviation == "PSMA") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "47.2"))

vtm_psma$dbh_class <- as.numeric(vtm_psma$dbh_class)
vtm_psma <- mutate(vtm_psma, canopy_width = (psma_a1 + psma_a2 * (dbh_class)))
vtm_psma$canopy_width <- as.numeric(vtm_psma$canopy_width)

## ------------------------------------------------------------------------------------------------# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_psma_canopy <- vtm_psma %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_psma_sum <- vtm_psma_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# QUCH Canopy Cover
```{r}
# Use Equation  4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)

vtm_quch <- vtm_data %>%
  filter(abbreviation == "QUCH") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count") %>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "36.2"))

vtm_quch$dbh_class <- as.numeric(vtm_quch$dbh_class)
vtm_quch <- mutate(vtm_quch, canopy_width = (quch_a1 + quch_a2 * (dbh_class)))
vtm_quch$canopy_width <- as.numeric(vtm_quch$canopy_width)

## ------------------------------------------------------------------------------------------------# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_quch_canopy <- vtm_quch %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_quch_sum <- vtm_quch_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# QUKE Canopy Cover
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)

vtm_quke <- vtm_data %>%
  filter(abbreviation == "QUKE") %>%
  select(plotkey, abbreviation,
         dbh_4_11, dbh_12_23, dbh_24_35, dbh_36,
         national_forest_name,
         elevation, exposure, slope_percent) %>%
  pivot_longer('dbh_4_11':'dbh_36',
               names_to = "dbh_class",
               values_to = "dbh_count")%>%
  filter(dbh_count !=0) %>%
  uncount(dbh_count) %>%
  mutate(dbh_class = case_when(
    dbh_class == "dbh_4_11" ~ "7.5",
    dbh_class == "dbh_12_23" ~ "17.5",
    dbh_class == "dbh_24_35" ~ "29.5",
    dbh_class == "dbh_36" ~ "36.6"))

vtm_quke$dbh_class <- as.numeric(vtm_quke$dbh_class)
vtm_quke <- mutate(vtm_quke, canopy_width = (quke_a1 + quke_a2 * (dbh_class)))
vtm_quke$canopy_width <- as.numeric(vtm_quke$canopy_width)

## ------------------------------------------------------------------------------------------------# Find canopy area based off of the known canopy widths. Canopy area =  the area of a circle with diameter equal to the determined crown width. The area of a circle is A = (pi)* r^2. 

vtm_quke_canopy <- vtm_quke %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2)) 
# canopy_area is now in feet^2

## ------------------------------------------------------------------------------------------------
# sum canopy cover across all size classes for each plot
vtm_quke_sum <- vtm_quke_canopy %>%
  group_by(plotkey, dbh_class, abbreviation, national_forest_name) %>%
  tally(canopy_area) 
```

# Combine all dataframes to calculate relative canopy cover
```{r}
# combine all species dataframes back together so we can sum the species in plots
vtm_total_canopy <- do.call("rbind", list(vtm_abco_sum,
                                          vtm_cade_sum,
                                          vtm_pico_sum,
                                          vtm_pije_sum,
                                          vtm_pila_sum,
                                          vtm_pipo_sum,
                                          vtm_psma_sum,
                                          vtm_quch_sum,
                                          vtm_quke_sum)) 
# n = the area (ft^2) covered by the canopies of all the trees in that size class/species/plot
## ------------------------------------------------------------------------------------------------
# Leana added this part
# Next we can sum the total canopy cover of all trees from all size classes PER PLOT
vtm_total_canopy_plot <- vtm_total_canopy %>% 
  group_by(plotkey, national_forest_name) %>% 
  tally(n) %>% 
  rename("total_canopy_area" = "n" )
# this checks out! There are 195 unique vtm plots! Leana checked a few plots and it looks good!
## ------------------------------------------------------------------------------------------------
# Leana added this part
# right now the total_canopy_area column is in ft^2, we want to convert this to m^2. 
# 1 ft^2 = 0.092903 m^2

# create a column that converts squared feet to squared meters
vtm_total_canopy_plot <- vtm_total_canopy_plot %>% 
  mutate(total_canopy_area_meters = (total_canopy_area *0.092903))

# Divide the total canopy area per plot (measured in m^2) by the size of the plot (also measured in m^2). Note: A vtm plot is 809 m^2. Then multiply by 100 to obtain a percent canopy cover. 
# Hannah: an FIA plot is 672.45 m^2.
vtm_total_canopy_plot <- vtm_total_canopy_plot %>% 
  mutate(percent_canopy_cover = ((total_canopy_area_meters/809) * 100))
# note that some of the percent covers are above 100%, because we have not yet accounted for overlapping canopies. 
# we have now determined the percent canopy cover without accounting for overlap--> equivalent to equation 1 in Crookston and Stage 1999 (https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf).
```

# Now apply equation 2 from Crookston and Stage 1999 to account for overlapping canopies. 
```{r}
# Equation 2 (page 2 of https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf):
# C = 100 [1 – exp ( – .01 C′ )]
# C = percent canopy cover that accounts for overlap,
# C′ = equation 1-percent canopy cover without accounting for overlap. *This has been calculated in the section above. 
# exp = e^

vtm_total_canopy_plot <- vtm_total_canopy_plot %>% 
  mutate(percent_canopy_cover_nonoverlap = (100*(1-(e^(-.01*percent_canopy_cover)))))
# I THINK it's okay to do this step here (after I've summed the canopy cover per plot), but I can go back and check if applying this equation earlier on changes our results. 
```

# Summarize findings
```{r}
# with canopy overlap
# average percent canopy cover across all YPMC plots when allowing for canopy overlap
mean(vtm_total_canopy_plot$percent_canopy_cover)
# 100.6753

canopy_cover_overlap <- vtm_total_canopy_plot %>% 
  mutate (national_forest_name = "All Plots") %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

canopy_cover_overlap_forests <- vtm_total_canopy_plot %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

canopy_cover_overlap_summary <- do.call("rbind", list(canopy_cover_overlap,
                                          canopy_cover_overlap_forests))

## -----------------------------------------------------------------------------------------------
# without canopy overlap
# average percent canopy cover across all YPMC plots when removing for canopy overlap
mean(vtm_total_canopy_plot$percent_canopy_cover_nonoverlap)
# 54.9521

canopy_cover_nonoverlap <- vtm_total_canopy_plot %>% 
  mutate (national_forest_name = "All Plots") %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

canopy_cover_nonoverlap_forests <- vtm_total_canopy_plot %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

canopy_cover_nonoverlap_summary <- do.call("rbind", list(canopy_cover_nonoverlap,
                                          canopy_cover_nonoverlap_forests))
```

# Make graphs
```{r}
# create final dataframe for graph
canopy_cover_nonoverlap_summary_rm_NA <-canopy_cover_nonoverlap_summary %>% 
  mutate(color = case_when(national_forest_name == "All Plots" ~ "1",
                           national_forest_name == "anf" ~ "0",
                           national_forest_name == "cnf" ~ "0",
                           national_forest_name == "lpnf" ~ "0",
                           national_forest_name == "sbnf" ~ "0",
                           national_forest_name == "N/A" ~ "0",
                           TRUE ~ "none"))
canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "anf"] <-"ANF"
canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "cnf"] <-"CNF"
canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "lpnf"] <-"LPNF"
canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "sbnf"] <-"SBNF"
canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "N/A"] <-"Other"
canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "All Plots"] <-"All YPMC"

# make factors
x_factor <- c("ANF", "CNF", "LPNF", "SBNF","Other", "All YPMC")
color_factor <- c("0", "1")

vtm_graph <-ggplot(canopy_cover_nonoverlap_summary_rm_NA,aes(x=factor(national_forest_name, x_factor),y=mean, fill=factor(color, color_factor)))+
   scale_fill_manual(values = c("#636161", "#A00D35"))+
  geom_col(stat="identity",position=position_dodge(width = 0.4), width = 0.4)+
  geom_point(data = canopy_cover_nonoverlap_summary_rm_NA, aes(x=national_forest_name, y= mean),position=position_dodge(width = 0.4), color = "black", size = 1.3, shape = 16)+
geom_errorbar(data = canopy_cover_nonoverlap_summary_rm_NA,
              position=position_dodge(width = 0.4),
                aes(x = national_forest_name, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC", x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0, 70))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.position = "none")

vtm_graph
```


##---------------------------------------------------------------------------------------------
### II. FIA
```{r}
# read in data & filter for dbh > 4 inches (10.16 cm)
fia_data <- read_csv(here::here("FIA Data","Final FIA data and code used for analysis",  "final_binded_fia_ypmc_data.csv")) %>%
  filter(dbh >= 10.2) # I'm assuming we still want to do this step because of VTM lacking this data?

# we also want to merge LPF into LPNF and BDF into SBNF
fia_data$admin_forest[fia_data$admin_forest == "LPF"] <-"LPNF"
fia_data$admin_forest[fia_data$admin_forest == "BDF"] <-"SBNF"
fia_data$admin_forest[fia_data$admin_forest == "Not NF or other NF"] <- "Other"
```

### Canopy Cover Overview
## Equations

# 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
# 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
# 4.4.1.4: CW = HT * s1 (HT <4.5' and DBH<DBHT)
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT)

# CW is maximum tree crown width
# DBH is tree diameter at breast height
# DBH(subT) is threshold diamter 

## Table (from...) Coefficients and equation reference for crown width equations {4.4.2.1}-{4.4.2.5} in the WS variant

# Sugar Pine (PILA), Equation 4.4.2.1, DBH(subT) = 7.4
# White fir (ABCO), Equation 4.4.2.1, DBH(subT) = 5
# Incense cedar (CADE), Equation 4.4.2.1, DBH(subT) = 5
# Jeffrey Pine (PIJE), Equation 4.4.2.2, DBH(subT) = 5
# Ponderosa Pine (PIPO), Equation 4.4.2.2, DBH(subT) = 5
# Coulter Pine (PICO), Equation 4.4.2.2, DBH(subT) = 5
# Bigcone Spruce (PSMA), Equation 4.4.2.1, DBH (subT) = 5
# Canyon Live Oak (QUCH), Equation 4.4.2.1, DBH(subT) = 5 
# Black Oak (QUKE), Equation 4.4.2.1, DBH(subT) = 5

```{r}
## Constants for equations 
# --------------------------------------------------
# Sugar Pine (PILA) constants:
pila_d1 <- 3.5
pila_d2 <- 0.338
pila_a1 <- -1.476
pila_a2 <- 1.01
pila_s1 <- 0.7778

# White fir (ABCO) constants:
abco_d1 <- 3.26
abco_d2 <- 1.103
abco_a1 <- 5.82
abco_a2 <- 0.591
abco_s1 <- 0.7778

# Incense Cedar (CADE) constants:
cade_d1 <- 3.5
cade_d2 <- 1.192
cade_a1 <- 7.11
cade_a2 <- 0.47
cade_s1 <- 0.7778

# Jeffrey Pine (PIJE) constants: 
pije_d1 <- 3.5
pije_d2 <- 0.5754
pije_a1 <- 1.52
pije_a2 <- 0.891
pije_s1 <- 0.7778

# Ponderosa Pine (PIPO) constants:
pipo_d1 <-3.77
pipo_d2 <- 0.7756
pipo_a1 <- 2.24
pipo_a2 <- 0.763
pipo_s1 <- 0.7778

# Coulter Pine (PICO) constants:
pico_d1 <- 3.5
pico_d2 <- 1.7618
pico_a1 <- 3.9347
pico_a2 <- 0.7086
pico_s1 <- 0.7778

# Bigcone Spruce (PSMA) constants
psma_d1 <- 3.5
psma_d2 <- 5.5672
psma_a1 <- 27.030
psma_a2 <- 0.8612
psma_s1 <- 0.7778

# Canyon Live Oak (QUCH) constants:
quch_d1 <- 2.5
quch_d2 <- 2.19
quch_a1 <- 5
quch_a2 <- 1.69
quch_s1 <- 0.5556

# Black Oak (QUKE) constants:
quke_d1 <- 2.5
quke_d2 <- 2.7
quke_a1 <- 10
quke_a2 <- 1.2
quke_s1 <- 0.5556

pi <- 3.1415926535 
e <- 2.718281828
```

```{r}
## Anne-Marie's midpoints for 36+ dbh
# vtm_basal_largest <- vtm_tidy %>% 
#   filter(dbh_class == "dbh_36") %>% 
#   mutate(dbh_final = case_when(
#     species == "ABCO" ~ 106 + 14,-->47.2
#     species == "CADE" ~ 162,-->63.8
#     species == "PICO" ~ 99 + 5,-->40.9
#     #species == "PIJE" ~ 104 + 12,-->45.7
#     species == "PILA" ~ 108 + 15,-->48.4
#     #species == "PIPO" ~ 110 + 18,-->50.4
#     species == "Yellow Pine" ~ 104 + 13,-->46.1
#     species == "PSMA" ~ 107 + 13,-->47.2
#     species == "QUCH" ~ 98 - 6,-->36.2
#     species == "QUKE" ~ 93-->36.6
```


## FIA Canopy Cover using midpoint of dbh class (same method as VTM)

# PIPO Canopy Cover (using midpoint of dbh bins)
```{r}
# Use Equation 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
# DBH(subT) is 5 and our lowest dbh is 7.5 since we are taking the midpoint of each class. 

# Select only PIPO & select only needed columns
# DBH for FIA is in centimeters! Still used the same inches midpoint that was used in vtm to be consistent. 

fia_pipo <- fia_data %>%
  filter(species == "PIPO") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "46.1")) # used Anne-Marie's measurements. Used the yellow pine value for pipo and pije, but we should double check this!

# change dbh_class to numeric (note, anything greater than 36 dbh won't have a plus or else they would change to NAs when converted to numeric)
fia_pipo$dbh_class <- as.numeric(fia_pipo$dbh_class)
# str(fia_pipo)
fia_pipo <- mutate(fia_pipo, canopy_width = (pipo_a1 * (dbh_class^(pipo_a2))))
fia_pipo$canopy_width <- as.numeric(fia_pipo$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_pipo_canopy <- fia_pipo %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_pipo_sum <- fia_pipo_canopy %>%
  group_by(unique_plot, dbh_class, admin_forest, species) %>%
  tally(canopy_area) # checked a couple of plots against the fia_pipo dataframe and they added up to the correct amount.
```

# PIJE Canopy Cover (using midpoint of dbh bins)
```{r}
# Equation for PIJE: # 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
fia_pije <- fia_data %>%
  filter(species == "PIJE") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "46.1")) # used Anne-Marie's measurements. Used the yellow pine value for pipo and pije, but we should double check this!

fia_pije$dbh_class <- as.numeric(fia_pije$dbh_class)
fia_pije <- mutate(fia_pije, canopy_width = (pije_a1 * (dbh_class^(pije_a2))))
fia_pije$canopy_width <- as.numeric(fia_pije$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_pije_canopy <- fia_pije %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_pije_sum <- fia_pije_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# PILA Canopy Cover (using midpoint of dbh class)
```{r}
# Use Equation: # 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
fia_pila <- fia_data %>%
  filter(species == "PILA") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "48.4")) 

fia_pila$dbh_class <- as.numeric(fia_pila$dbh_class)
fia_pila <- mutate(fia_pila, canopy_width = (pila_a1 + pila_a2 * (dbh_class))) 
fia_pila$canopy_width <- as.numeric(fia_pila$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_pila_canopy <- fia_pila %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_pila_sum <- fia_pila_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# PICO Canopy Cover (using midpoint dbh class)
```{r}
# Use Equation 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT)
fia_pico <- fia_data %>%
  filter(species == "PICO") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "40.9")) 

fia_pico$dbh_class <- as.numeric(fia_pico$dbh_class)
fia_pico <- mutate(fia_pico, canopy_width = (pico_a1 * (dbh_class^(pico_a2))))
fia_pico$canopy_width <- as.numeric(fia_pico$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_pico_canopy <- fia_pico %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_pico_sum <- fia_pico_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# ABCO Canopy Cover (using midpoint dbh class)
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
fia_abco <- fia_data %>%
  filter(species == "ABCO") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "47.2")) 

fia_abco$dbh_class <- as.numeric(fia_abco$dbh_class)
fia_abco <- mutate(fia_abco, canopy_width = (abco_a1 + abco_a2 * (dbh_class)))
fia_abco$canopy_width <- as.numeric(fia_abco$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_abco_canopy <- fia_abco %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_abco_sum <- fia_abco_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# CADE Canopy Cover (using midpoint dbh class)
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
fia_cade <- fia_data %>%
  filter(species == "CADE") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "63.8")) 

fia_cade$dbh_class <- as.numeric(fia_cade$dbh_class)
fia_cade <- mutate(fia_cade, canopy_width = (cade_a1 + cade_a2 * (dbh_class)))
fia_cade$canopy_width <- as.numeric(fia_cade$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_cade_canopy <- fia_cade %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_cade_sum <- fia_cade_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# PSMA Canopy Cover (using midpoint dbh class)
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
fia_psma <- fia_data %>%
  filter(species == "PSMA") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "47.2")) 

fia_psma$dbh_class <- as.numeric(fia_psma$dbh_class)
fia_psma <- mutate(fia_psma, canopy_width = (psma_a1 + psma_a2 * (dbh_class)))
fia_psma$canopy_width <- as.numeric(fia_psma$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_psma_canopy <- fia_psma %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_psma_sum <- fia_psma_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# QUCH Canopy Cover (using midpoint dbh class)
```{r}
# Use Equation  4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
fia_quch <- fia_data %>%
  filter(species == "QUCH") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "36.2")) 

fia_quch$dbh_class <- as.numeric(fia_quch$dbh_class)
fia_quch <- mutate(fia_quch, canopy_width = (quch_a1 + quch_a2 * (dbh_class)))
fia_quch$canopy_width <- as.numeric(fia_quch$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_quch_canopy <- fia_quch %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_quch_sum <- fia_quch_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# QUKE Canopy Cover (using midpoint dbh class)
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT)
fia_quke <- fia_data %>%
  filter(species == "QUKE") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_class = case_when (dbh >= 10.2 & dbh <= 30.4 ~"7.5",
                                dbh > 30.4 & dbh <= 60.9 ~ "17.5",
                                dbh > 60.9 & dbh <= 91.3 ~ "29.5",
                                dbh > 91.3 ~ "36.6")) 

fia_quke$dbh_class <- as.numeric(fia_quke$dbh_class)
fia_quke <- mutate(fia_quke, canopy_width = (quke_a1 + quke_a2 * (dbh_class)))
fia_quke$canopy_width <- as.numeric(fia_quke$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_quke_canopy <- fia_quke %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_quke_sum <- fia_quke_canopy %>%
  group_by(unique_plot, dbh_class, species, admin_forest) %>%
  tally(canopy_area) 
```

# Combine all dataframes to calculate relative canopy cover
```{r}
# combine all species dataframes back together so we can sum the species in plots
fia_total_canopy <- do.call("rbind", list(fia_abco_sum,
                                          fia_cade_sum,
                                          fia_pico_sum,
                                          fia_pije_sum,
                                          fia_pila_sum,
                                          fia_pipo_sum,
                                          fia_psma_sum,
                                          fia_quch_sum,
                                          fia_quke_sum))

## -------------------------------------------------------------------------------------------------
# Next we can sum the total canopy cover of all trees from all size classes PER PLOT
fia_total_canopy_plot <- fia_total_canopy %>% 
  group_by(unique_plot, admin_forest) %>% 
  tally(n) %>% 
  rename("total_canopy_area" = "n" )

## ------------------------------------------------------------------------------------------------
# right now the total_canopy_area column is in ft^2, we want to convert this to m^2. 
# 1 ft^2 = 0.092903 m^2

# create a column that converts squared feet to squared meters
fia_total_canopy_plot <- fia_total_canopy_plot %>% 
  mutate(total_canopy_area_meters = (total_canopy_area *0.092903))

# Divide the total canopy area per plot (measured in m^2) by the size of the plot (also measured in m^2). Note: A vtm plot is 672.45 m^2. Then multiply by 100 to obtain a percent canopy cover. 
fia_total_canopy_plot <- fia_total_canopy_plot %>% 
  mutate(percent_canopy_cover = ((total_canopy_area_meters/672.45) * 100))

# note that some of the percent covers are above 100%, because we have not yet accounted for overlapping canopies. 
# we have now determined the percent canopy cover without accounting for overlap--> equivalent to equation 1 in Crookston and Stage 1999 (https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf).
```

# Now apply equation 2 from Crookston and Stage 1999 to account for overlapping canopies. 
```{r}
# Equation 2 (page 2 of https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf):
# C = 100 [1 – exp ( – .01 C′ )]
# C = percent canopy cover that accounts for overlap,
# C′ = equation 1-percent canopy cover without accounting for overlap. *This has been calculated in the section above. 
# exp = e^

fia_total_canopy_plot <- fia_total_canopy_plot %>% 
  mutate(percent_canopy_cover_nonoverlap = (100*(1-(e^(-.01*percent_canopy_cover)))))
# I THINK it's okay to do this step here (after I've summed the canopy cover per plot), but I can go back and check if applying this equation earlier on changes our results. 
```

# Summarize Findings
```{r}
# with canopy overlap
# average percent canopy cover across all YPMC plots when allowing for canopy overlap
mean(fia_total_canopy_plot$percent_canopy_cover)
# 167.2744

fia_canopy_cover_overlap <- fia_total_canopy_plot %>% 
  mutate (admin_forest = "All Plots") %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

fia_canopy_cover_overlap_forests <- fia_total_canopy_plot %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

fia_canopy_cover_overlap_summary <- do.call("rbind", list(fia_canopy_cover_overlap,
                                          fia_canopy_cover_overlap_forests))

## -----------------------------------------------------------------------------------------------
# without canopy overlap
# average percent canopy cover across all YPMC plots when removing for canopy overlap
mean(fia_total_canopy_plot$percent_canopy_cover_nonoverlap)
# 69.93036

fia_canopy_cover_nonoverlap <- fia_total_canopy_plot %>% 
  mutate (admin_forest = "All Plots") %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

fia_canopy_cover_nonoverlap_forests <- fia_total_canopy_plot %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

fia_canopy_cover_nonoverlap_summary <- do.call("rbind", list(fia_canopy_cover_nonoverlap,
                                          fia_canopy_cover_nonoverlap_forests))
```

# Make Graph 
```{r}
fia_canopy_cover_nonoverlap_summary_rm_NA <-fia_canopy_cover_nonoverlap_summary %>% 
  mutate(color = case_when(admin_forest == "All Plots" ~ "1",
                           admin_forest == "ANF" ~ "0",
                           admin_forest == "CNF" ~ "0",
                           admin_forest == "LPNF" ~ "0",
                           admin_forest == "SBNF" ~ "0",
                           admin_forest == "Other" ~ "0",
                           TRUE ~ "none"))

# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "anf"] <-"ANF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "cnf"] <-"CNF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "lpnf"] <-"LPNF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "sbnf"] <-"SBNF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "N/A"] <-"Other"
fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "All Plots"] <-"All YPMC"

# make factors
x_factor <- c("ANF", "CNF", "LPNF", "SBNF","Other", "All YPMC")
color_factor <- c("0", "1")

fia_graph <-ggplot(fia_canopy_cover_nonoverlap_summary_rm_NA,aes(x=factor(admin_forest, x_factor),y=mean, fill=factor(color, color_factor)))+
   scale_fill_manual(values = c("#636161", "#A00D35"))+
  geom_col(stat="identity",position=position_dodge(width = 0.4), width = 0.4)+
  geom_point(data = fia_canopy_cover_nonoverlap_summary_rm_NA, aes(x=admin_forest, y= mean),position=position_dodge(width = 0.4), color = "black", size = 1.3, shape = 16)+
geom_errorbar(data = fia_canopy_cover_nonoverlap_summary_rm_NA,
              position=position_dodge(width = 0.4),
                aes(x = admin_forest, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC",
       x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0, 90))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.position = "none") 

fia_graph
```

### III. Comparison
```{r}
#combine datasets
#canopy_cover_nonoverlap_summary_rm_NA vtm data
#canopy_cover_nonoverlap_summary_rm_NA fia data

fia_canopy_cover_nonoverlap_summary_rm_NA <- fia_canopy_cover_nonoverlap_summary_rm_NA %>%
  filter(admin_forest %in% c("ANF", "SBNF", "LPNF", "Other", "All YPMC"))

canopy_cover_nonoverlap_summary_rm_NA_2 <- canopy_cover_nonoverlap_summary_rm_NA %>%
   filter(national_forest_name %in% c("ANF", "SBNF", "LPNF", "Other", "All YPMC"))

canopy_cover_nonoverlap_summary_rm_NA_2 <- canopy_cover_nonoverlap_summary_rm_NA %>% 
  mutate(plot = case_when(national_forest_name == "ANF" ~ "VTM",
                          national_forest_name == "SBNF" ~ "VTM",
                          national_forest_name == "LPNF" ~ "VTM",
                          national_forest_name == "Other" ~ "VTM",
                          national_forest_name == "All YPMC" ~ "VTM_1",
                          TRUE ~ "none"))
  

fia_canopy_cover_nonoverlap_summary_rm_NA_2 <- fia_canopy_cover_nonoverlap_summary_rm_NA %>% 
  mutate(plot = case_when(admin_forest == "ANF" ~ "FIA",
                          admin_forest == "SBNF" ~ "FIA",
                          admin_forest == "LPNF" ~ "FIA",
                          admin_forest == "Other" ~ "FIA",
                          admin_forest == "All YPMC" ~ "FIA_1",
                          TRUE ~ "none")) %>% 
  rename( "national_forest_name" = "admin_forest")


canopy_cover_comparison <- do.call("rbind", list(canopy_cover_nonoverlap_summary_rm_NA_2,
                                          fia_canopy_cover_nonoverlap_summary_rm_NA_2))

canopy_cover_comparison_2 <- canopy_cover_comparison %>% 
  filter(national_forest_name == "All YPMC")

canopy_cover_comparison <- canopy_cover_comparison %>%
  filter(national_forest_name %in% c("ANF", "SBNF", "LPNF", "Other", "All YPMC"))

```


# Percent change
```{r}
pct_change_binned_fia <- canopy_cover_comparison %>%
  select(national_forest_name, mean, plot) %>%
  group_by(national_forest_name) %>%
  mutate(pct_change = ((mean/lag(mean) - 1)*100))
```

# make comparison graph
```{r}
x_factor <- c("ANF", "LPNF", "SBNF","Other", "All YPMC")
color_factor <- c("0", "1")
order_factor <- c("VTM", "FIA", "VTM_1", "FIA_1")

ggplot(canopy_cover_comparison,aes(x=factor(national_forest_name, x_factor),y=mean, fill=factor(plot, order_factor)))+
   scale_fill_manual(values = c("#9C9A9A","#636161", "#BD606E", "#A00D35"), labels=c("VTM", "FIA", "VTM", "FIA"))+
  geom_col(stat="identity",position=position_dodge(width = 0.7), width = 0.7)+
    geom_text(data = canopy_cover_comparison_2, aes(label=round(mean, 1)), position=position_dodge(width = 0.7), vjust=-1.6, size = 4)+
  geom_point(data = canopy_cover_comparison, aes(x=national_forest_name, y= mean),position=position_dodge(width = 0.7), color = "black", size = 1.3, shape = 16, show.legend = FALSE)+
geom_errorbar(data = canopy_cover_comparison,
              position=position_dodge(width = 0.7),
                aes(x = national_forest_name, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC (Including Oaks)",
       subtitle = "Using mean + 1SD for the 36+ size class midpoint", 
       x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(plot.subtitle = element_text(hjust= 0.5,
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0, 90))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.title = element_blank()) +
  ggsave(here::here("NRV Analysis", "canopy_cover", "Figures", "vtm_fia_canopycover_mean+1sd_midpoint_withoaks.png"),
         height=5,
         width=7,
         units="in")
```

### IV. Compare binned and unbinned fia dbh measurements

## FIA Canopy Cover using actual dbh measurement 
## FIA height column is in meters
```{r}
# convert height column from meters to ft for the equations
fia_data <- read_csv(here::here("FIA Data", "Final FIA data and code used for analysis", "final_binded_fia_ypmc_data.csv")) %>%
  filter(dbh >= 10.2) %>%
  mutate(ht_ft = ht * 3.28084)

fia_data$admin_forest[fia_data$admin_forest == "LPF"] <-"LPNF"
fia_data$admin_forest[fia_data$admin_forest == "BDF"] <-"SBNF"
fia_data$admin_forest[fia_data$admin_forest == "Not NF or other NF"] <- "Other"
```

# PIPO Canopy cover (final dataframe = fia_pipo_two_sum) -- done
```{r}
# Use Equation 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT) DBH(subT) = 5
fia_pipo_two <- fia_data %>%
  filter(species == "PIPO") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54)) 

# check to see if any under 5 (which is the threshold)
# fia_pipo_two_check <- fia_pipo_two %>%
#   filter(dbh_in < 5) # none under 5, which means we can use the above equation

fia_pipo_two$dbh_in <- as.numeric(fia_pipo_two$dbh_in )
# calculate canopy cover
fia_pipo_two <- mutate(fia_pipo_two, canopy_width = (pipo_a1 * (dbh_in^(pipo_a2))))
fia_pipo_two$canopy_width <- as.numeric(fia_pipo_two$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_pipo_two_canopy <- fia_pipo_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_pipo_two_sum <- fia_pipo_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)

# comments: since we are not lumping size classes, wonder if this last step even makes sense? whatever, it doesnt hurt.
```

# PIJE Canopy Cover (final dataframe = fia_pije_two_sum) -- done
```{r}
# Use Equation 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT) DBH(subT) = 5
fia_pije_two <- fia_data %>%
  filter(species == "PIJE") %>%
  select(unique_plot, dbh, ht_ft,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54)) 

# check to see if any under 5 (which is threshhold)
fia_pije_two_check <- fia_pije_two %>%
  filter(dbh_in < 4.9) # yes, there are 9 observations under 5

fia_pije_two$dbh_in <- as.numeric(fia_pije_two$dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh larger or equal to 5
fia_pije_greater_than5 <- fia_pije_two %>%
  filter(dbh_in >= 4.9) %>%
  mutate(canopy_width = pije_a1 * (dbh_in^pije_a2))

## -------------------------------------------------------------------------------------------------
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT) 
# all 9 observations taller than 4.5 but less than 5 dbh

fia_pije_less_than5_tall <- fia_pije_two_check %>%
  filter(dbh_in < 4.9) %>%
  mutate(canopy_width = pije_d1 + pije_d2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# combine two dataframes
pije_total_two <- do.call("rbind", list(fia_pije_greater_than5,
                                        fia_pije_less_than5_tall))

## -------------------------------------------------------------------------------------------------
fia_pije_two_canopy <- pije_total_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_pije_two_sum <- fia_pije_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)
```

# PILA Canopy Cover (final dataframe = fia_pila_two_sum) -- done
```{r}
# Use Equation: # 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT), DBH(subT) = 7.4
fia_pila_two <- fia_data %>%
  filter(species == "PILA") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54))

fia_pila_two_check <- fia_pila_two %>%
  filter(dbh_in < 7.4) # 21 trees under the threshold value, no trees under 4.5' in height

fia_pila_two$dbh_in <- as.numeric(fia_pila_two$dbh_in)
## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh larger or equal to 7.4
fia_pila_greater_than7 <- fia_pila_two %>%
  filter(dbh_in >= 7.4) %>%
  mutate(canopy_width = pila_a1 + pila_a2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh less than 7.4 but taller than 4.5 inches
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT)
fia_pila_less_than7 <- fia_pila_two %>%
  filter(dbh_in < 7.4) %>%
  mutate(canopy_width = pila_d1 + pila_d2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# Combine two dataframes
pila_total_two <- do.call("rbind", list(fia_pila_greater_than7,
                                        fia_pila_less_than7))

## -------------------------------------------------------------------------------------------------
fia_pila_two_canopy <- pila_total_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes for each plot
fia_pila_two_sum <- fia_pila_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)
```

# PICO Canopy Cover (final dataframe = fia_pico_two_sum) -- done
```{r}
# Use Equation 4.4.2.2: CW = a1 * DBH^a2 (DBH>=DBHT), DBH(subT) = 5
fia_pico_two <- fia_data %>%
  filter(species == "PICO") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54))

fia_pico_two_check <- fia_pico_two %>%
  filter(dbh_in < 4.9) # only one tree under the threshold dbh and is taller than 4.5 ft

fia_pico_two$dbh_in <- as.numeric(fia_pico_two$dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh larger or equal to 5
fia_pico_greater_than5 <- fia_pico_two %>%
  filter(dbh_in >= 4.9) %>%
  mutate(canopy_width = pico_a1 * (dbh_in^pico_a2))

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh less than 5, ht greater than 4.5 in
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT)
fia_pico_less_than5 <- fia_pico_two %>%
  filter(dbh_in < 5) %>%
   mutate(canopy_width = pico_d1 + pico_d2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# combine the two dataframes
pico_total_two <- do.call("rbind", list(fia_pico_less_than5,
                                        fia_pico_greater_than5))

## -------------------------------------------------------------------------------------------------
fia_pico_two_canopy <- pico_total_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# Sum across size classes
fia_pico_two_sum <- fia_pico_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)

```

# ABCO Canopy Cover (final dataframe = fia_abco_two_sum) -- done
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT), DBH(subT) = 5
fia_abco_two <- fia_data %>%
  filter(species == "ABCO") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54))

fia_abco_two_check <- fia_abco_two %>%
  filter(dbh_in < 4.9) # 8 trees with dbh under the threshold level and 5 under 4.5 ft

fia_abco_two$dbh_in <- as.numeric(fia_abco_two$dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh larger or equal to 5
fia_abco_greater_than5 <- fia_abco_two %>%
  filter(dbh_in >= 4.9) %>%
  mutate(canopy_width = abco_a1 + abco_a2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh less than 5, with a height greather than 4.5
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT)
fia_abco_less_than5_tall <- fia_abco_two_check %>%
  filter(ht > 4.5) %>%
  mutate(canopy_width = abco_d1 + abco_d2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh less than 5, with a height less than 4.5
## 4.4.1.4: CW = HT * s1 (HT <4.5' and DBH<DBHT)
fia_abco_less_than5_short <- fia_abco_two_check %>%
  filter(ht < 4.5) %>%
  mutate(canopy_width = ht * abco_s1)

## -------------------------------------------------------------------------------------------------
# combine the three dataframes
abco_total_two <- do.call("rbind", list(fia_abco_less_than5_short,
                                        fia_abco_less_than5_tall,
                                        fia_abco_greater_than5))

## -------------------------------------------------------------------------------------------------
fia_abco_two_canopy <- abco_total_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# Sum across size classes
fia_abco_two_sum <- fia_abco_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)
```

# CADE Canopy Cover (final dataframe = fia_cade_two_sum) -- done
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT), DBH(subT) = 5
fia_cade_two <- fia_data %>%
  filter(species == "CADE") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54))

fia_cade_two_check <- fia_cade_two %>%
  filter(dbh_in < 4.9) # No trees under the threshold dbh WOO

fia_cade_two$dbh_in <- as.numeric(fia_cade_two$dbh_in)
fia_cade_two <- mutate(fia_cade_two, canopy_width = (cade_a1 + cade_a2 * (dbh_in)))
fia_cade_two$canopy_width <- as.numeric(fia_cade_two$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_cade_two_canopy <- fia_cade_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# Sum across size classes
fia_cade_two_sum <- fia_cade_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)
```

# PSMA Canopy Cover (final dataframe = fia_psma_two_sum) -- done
```{r}
# Use Equation 4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT), DBH (subT) = 5
fia_psma_two <- fia_data %>%
  filter(species == "PSMA") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54))

fia_psma_two_check <- fia_psma_two %>%
  filter(dbh_in < 4.9) # No trees under threshold :)

fia_psma_two$dbh_in <- as.numeric(fia_psma_two$dbh_in)
fia_psma_two <- mutate(fia_psma_two, canopy_width = (psma_a1 + psma_a2 * (dbh_in)))
fia_psma_two$canopy_width <- as.numeric(fia_psma_two$canopy_width)

## -------------------------------------------------------------------------------------------------
fia_psma_two_canopy <- fia_psma_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# Sum across size classes
fia_psma_two_sum <- fia_psma_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)
```

# QUCH Canopy Cover (final dataframe = fia_quch_two_sum) -- done
```{r}
# Use Equation  4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT), DBH(subT) = 5 
fia_quch_two <- fia_data %>%
  filter(species == "QUCH") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54))

fia_quch_two_check <- fia_quch_two %>%
   filter(dbh_in < 4.9) # 44 under the threshold

fia_quch_two$dbh_in <- as.numeric(fia_quch_two$dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for all trees with dbh larger or equal to 5
fia_quch_greater_than5 <- fia_quch_two %>%
  filter(dbh_in > 4.9) %>%
  mutate(canopy_width = quch_a1 + quch_a2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh less than 5 but taller than 4.5 (21 trees)
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT)
fia_quch_less_than5_tall <- fia_quch_two_check %>%
  filter(ht > 4.5) %>%
  mutate(canopy_width = quch_d1 + quch_d2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for trees with dbh less than 5 but shorter than 4.5 (23 trees)
# 4.4.1.4: CW = HT * s1 (HT <4.5' and DBH<DBHT)
fia_quch_less_than5_short <- fia_quch_two_check %>%
  filter(ht < 4.5) %>%
  mutate(canopy_width = ht * quch_s1)

## -------------------------------------------------------------------------------------------------
# combine the three dataframes
quch_total_two <- do.call("rbind", list(fia_quch_greater_than5,
                                       fia_quch_less_than5_tall,
                                       fia_quch_less_than5_short))

## -------------------------------------------------------------------------------------------------
fia_quch_two_canopy <- quch_total_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# Sum across size classes
fia_quch_two_sum <- fia_quch_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)
```

# QUKE Canopy Cover (final dataframe = fia_quke_two_sum) -- done
```{r}
# Use Equation  4.4.2.1: CW = a1 + a2 * DBH (DBH>=DBHT), DBH(subT) = 5
fia_quke_two <- fia_data %>%
  filter(species == "QUKE") %>%
  select(unique_plot, dbh, ht,
         admin_forest, species) %>%
  mutate(dbh_in = dbh * (1/2.54))

fia_quke_two_check <- fia_quke_two %>%
   filter(dbh_in < 4.9) # 1 tree under threshold

fia_quke_two$dbh_in <- as.numeric(fia_quke_two$dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for all trees over dbh of 5
fia_quke_greater_than5 <- fia_quke_two %>%
  filter(dbh_in > 4.9) %>%
  mutate(canopy_width = quke_a1 + quke_a2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# canopy cover for all trees with dbh less than 5, ht greater than 4.5 (1 tree)
# 4.4.1.5: CW = d1 + d2 * DBH (HT>4.5' and DBH<DBHT)
fia_quke_less_than5 <- fia_quke_two_check %>%
  mutate(canopy_width = quke_d1 + quke_d2 * dbh_in)

## -------------------------------------------------------------------------------------------------
# combine all dataframes
quke_total_two <- do.call("rbind", list(fia_quke_greater_than5,
                                        fia_quke_less_than5))

## -------------------------------------------------------------------------------------------------
fia_quke_two_canopy <- quke_total_two %>%
  mutate(canopy_area = (pi) * ((0.5 * canopy_width)^2))

## -------------------------------------------------------------------------------------------------
# sum across all size classes
fia_quke_two_sum <- fia_quke_two_canopy %>%
  group_by(unique_plot, dbh_in, admin_forest, species) %>%
  tally(canopy_area)
```

# combine all the dataframes from method 2 (no binned dbh)
```{r}
fia_total_canopy_two <- do.call("rbind", list(fia_quke_two_sum,
                                              fia_quch_two_sum,
                                              fia_psma_two_sum,
                                              fia_cade_two_sum,
                                              fia_abco_two_sum,
                                              fia_pico_two_sum,
                                              fia_pila_two_sum,
                                              fia_pije_two_sum,
                                              fia_pipo_two_sum))

## ------------------------------------------------------------------------------------------------
# Next we can sum the total canopy cover of all trees from all size classes PER PLOT
fia_total_canopy_plot_two <- fia_total_canopy_two %>% 
  group_by(unique_plot, admin_forest) %>% 
  tally(n) %>% 
  rename("total_canopy_area" = "n" )

## ------------------------------------------------------------------------------------------------
# right now the total_canopy_area column is in ft^2, we want to convert this to m^2. 
# 1 ft^2 = 0.092903 m^2

# create a column that converts squared feet to squared meters
fia_total_canopy_plot_two <- fia_total_canopy_plot_two %>% 
  mutate(total_canopy_area_meters = (total_canopy_area *0.092903))

# Divide the total canopy area per plot (measured in m^2) by the size of the plot (also measured in m^2). Note: A vtm plot is 672.45 m^2. Then multiply by 100 to obtain a percent canopy cover. 
fia_total_canopy_plot_two <- fia_total_canopy_plot_two %>% 
  mutate(percent_canopy_cover = ((total_canopy_area_meters/672.45) * 100))
# note that some of the percent covers are above 100%, because we have not yet accounted for overlapping canopies. 
# we have now determined the percent canopy cover without accounting for overlap--> equivalent to equation 1 in Crookston and Stage 1999 (https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf).
```

# Now apply equation 2 from Crookston and Stage 1999 to account for overlapping canopies. 
```{r}
# Equation 2 (page 2 of https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf):
# C = 100 [1 – exp ( – .01 C′ )]
# C = percent canopy cover that accounts for overlap,
# C′ = equation 1-percent canopy cover without accounting for overlap. *This has been calculated in the section above. 
# exp = e^

fia_total_canopy_plot_two <- fia_total_canopy_plot_two %>% 
  mutate(percent_canopy_cover_nonoverlap = (100*(1-(e^(-.01*percent_canopy_cover)))))
# I THINK it's okay to do this step here (after I've summed the canopy cover per plot), but I can go back and check if applying this equation earlier on changes our results. 
```

# Summarize Findings
```{r}
# with canopy overlap
# average percent canopy cover across all YPMC plots when allowing for canopy overlap
mean(fia_total_canopy_plot_two$percent_canopy_cover)
# 155.6348

canopy_cover_overlap_two <- fia_total_canopy_plot_two %>% 
  mutate (admin_forest = "All Plots") %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

canopy_cover_overlap_forests_two <- fia_total_canopy_plot_two %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

canopy_cover_overlap_summary_two <- do.call("rbind", list(canopy_cover_overlap_two,
                                          canopy_cover_overlap_forests_two))

## -----------------------------------------------------------------------------------------------
# without canopy overlap
# average percent canopy cover across all YPMC plots when removing for canopy overlap
mean(fia_total_canopy_plot_two$percent_canopy_cover_nonoverlap)
# 67.8215

canopy_cover_nonoverlap_two <- fia_total_canopy_plot_two %>% 
  mutate (admin_forest = "All Plots") %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

canopy_cover_nonoverlap_forests_two <- fia_total_canopy_plot_two %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

canopy_cover_nonoverlap_summary_two <- do.call("rbind", list(canopy_cover_nonoverlap_two,
                                          canopy_cover_nonoverlap_forests_two))
```

# Make Graphs
```{r}
# create final dataframe for graph
canopy_cover_nonoverlap_summary_rm_NA_two <-canopy_cover_nonoverlap_summary_two %>% 
  mutate(color = case_when(admin_forest == "All Plots" ~ "1",
                           admin_forest == "ANF" ~ "0",
                           admin_forest == "CNF" ~ "0",
                           admin_forest == "LPNF" ~ "0",
                           admin_forest == "SBNF" ~ "0",
                           admin_forest == "Other" ~ "0",
                           TRUE ~ "none"))
# canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "anf"] <-"ANF"
# canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "cnf"] <-"CNF"
# canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "lpnf"] <-"LPNF"
# canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "sbnf"] <-"SBNF"
# canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "N/A"] <-"Other"
# canopy_cover_nonoverlap_summary_rm_NA$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA$national_forest_name == "All Plots"] <-"All YPMC"

# make factors
x_factor <- c("ANF", "CNF", "LPNF", "SBNF","Other", "All Plots")
color_factor <- c("0", "1")

 ggplot(canopy_cover_nonoverlap_summary_rm_NA_two,aes(x=factor(admin_forest, x_factor),y=mean, fill=factor(color, color_factor)))+
   scale_fill_manual(values = c("#636161", "#A00D35"))+
  geom_col(stat="identity",position=position_dodge(width = 0.4), width = 0.4)+
  geom_point(data = canopy_cover_nonoverlap_summary_rm_NA_two, aes(x=admin_forest, y= mean),position=position_dodge(width = 0.4), color = "black", size = 1.3, shape = 16)+
geom_errorbar(data = canopy_cover_nonoverlap_summary_rm_NA_two,
              position=position_dodge(width = 0.4),
                aes(x = admin_forest, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC", subtitle = "Using acutal dbh measurements", x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(plot.subtitle = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0, 90))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.position = "none")
 
```

# combine the dataframes
```{r}
#combine datasets
# fia_canopy_cover_nonoverlap_summary_rm_NA fia data method one
# canopy_cover_nonoverlap_summary_rm_NA_two fia method two
  
fia_canopy_cover_nonoverlap_summary_rm_NA_2 <- fia_canopy_cover_nonoverlap_summary_rm_NA %>% 
  mutate(plot = case_when(admin_forest == "ANF" ~ "FIA_1",
                          admin_forest == "CNF" ~ "FIA_1",
                          admin_forest == "SBNF" ~ "FIA_1",
                          admin_forest == "LPNF" ~ "FIA_1",
                          admin_forest == "Other" ~ "FIA_1",
                          admin_forest == "All YPMC" ~ "FIA_1",
                          TRUE ~ "none")) 

fia_canopy_cover_nonoverlap_summary_rm_NA_3 <- canopy_cover_nonoverlap_summary_rm_NA_two %>% 
  mutate(plot = case_when(admin_forest == "ANF" ~ "FIA_2",
                          admin_forest == "CNF" ~ "FIA_2",
                          admin_forest == "SBNF" ~ "FIA_2",
                          admin_forest == "LPNF" ~ "FIA_2",
                          admin_forest == "Other" ~ "FIA_2",
                          admin_forest == "All YPMC" ~ "FIA_2",
                          TRUE ~ "FIA_2")) 
fia_canopy_cover_nonoverlap_summary_rm_NA_3$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA_3$admin_forest == "All Plots"] <-"All YPMC"

fia_canopy_cover_comparison_method <- do.call("rbind", list(fia_canopy_cover_nonoverlap_summary_rm_NA_2,
                                                            fia_canopy_cover_nonoverlap_summary_rm_NA_3))

fia_canopy_cover_comparison_method_2 <- fia_canopy_cover_comparison_method %>% 
  filter(admin_forest == "All YPMC")
```

# Make a graph
```{r}
x_factor <- c("ANF", "CNF", "LPNF", "SBNF","Other", "All YPMC")
color_factor <- c("0", "1")
order_factor <- c("FIA_1", "FIA_2")

ggplot(fia_canopy_cover_comparison_method, aes(x=factor(admin_forest, x_factor),y=mean, fill=factor(plot, order_factor)))+
   scale_fill_manual(values = c("#9C9A9A","#636161", "#BD606E", "#A00D35"), labels=c("FIA_1", "FIA_2"))+
  geom_col(stat="identity",position=position_dodge(width = 0.7), width = 0.7)+
    geom_text(data = fia_canopy_cover_comparison_method_2, aes(label=round(mean, 1)), position=position_dodge(width = 0.7), vjust=-1.6, size = 4)+
  geom_point(data = fia_canopy_cover_comparison_method, aes(x=admin_forest, y= mean),position=position_dodge(width = 0.7), color = "black", size = 1.3, shape = 16, show.legend = FALSE)+
geom_errorbar(data = fia_canopy_cover_comparison_method,
              position=position_dodge(width = 0.7),
                aes(x = admin_forest, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC", 
       subtitle = "Comparing Binned DBH and Actual DBH Measurements",
       x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(plot.subtitle= element_text(hjust= 0.5,
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0, 90))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.title = element_blank())
```

### V. Compare canopy cover without oaks (as recommended at fall faculty review)

## VTM (without oaks)
# Combine all dataframes to calculate relative canopy cover
```{r}
# combine all species dataframes back together so we can sum the species in plots
vtm_total_canopy_conifer <- do.call("rbind", list(vtm_abco_sum,
                                          vtm_cade_sum,
                                          vtm_pico_sum,
                                          vtm_pije_sum,
                                          vtm_pila_sum,
                                          vtm_pipo_sum,
                                          vtm_psma_sum)) 
# n = the area (ft^2) covered by the canopies of all the trees in that size class/species/plot
## -----------------------------------------------------------------------------------------------

# Next we can sum the total canopy cover of all trees from all size classes PER PLOT
vtm_total_canopy_plot_conifer <- vtm_total_canopy_conifer %>% 
  group_by(plotkey, national_forest_name) %>% 
  tally(n) %>% 
  rename("total_canopy_area" = "n" )
# this checks out! There are 195 unique vtm plots! Leana checked a few plots and it looks good!
## -----------------------------------------------------------------------------------------------

# right now the total_canopy_area column is in ft^2, we want to convert this to m^2. 
# 1 ft^2 = 0.092903 m^2

# create a column that converts squared feet to squared meters
vtm_total_canopy_plot_conifer <- vtm_total_canopy_plot_conifer %>% 
  mutate(total_canopy_area_meters = (total_canopy_area *0.092903))

# Divide the total canopy area per plot (measured in m^2) by the size of the plot (also measured in m^2). Note: A vtm plot is 809 m^2. Then multiply by 100 to obtain a percent canopy cover. 
# Hannah: an FIA plot is 672.45 m^2.
vtm_total_canopy_plot_conifer <- vtm_total_canopy_plot_conifer %>% 
  mutate(percent_canopy_cover = ((total_canopy_area_meters/809) * 100))
# note that some of the percent covers are above 100%, because we have not yet accounted for overlapping canopies. 
# we have now determined the percent canopy cover without accounting for overlap--> equivalent to equation 1 in Crookston and Stage 1999 (https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf).
```

# Now apply equation 2 from Crookston and Stage 1999 to account for overlapping canopies. 
```{r}
# Equation 2 (page 2 of https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf):
# C = 100 [1 – exp ( – .01 C′ )]
# C = percent canopy cover that accounts for overlap,
# C′ = equation 1-percent canopy cover without accounting for overlap. *This has been calculated in the section above. 
# exp = e^

vtm_total_canopy_plot_conifer <- vtm_total_canopy_plot_conifer %>% 
  mutate(percent_canopy_cover_nonoverlap = (100*(1-(e^(-.01*percent_canopy_cover)))))
# I THINK it's okay to do this step here (after I've summed the canopy cover per plot), but I can go back and check if applying this equation earlier on changes our results. 
```

# Summarize findings
```{r}
# with canopy overlap
# average percent canopy cover across all YPMC plots when allowing for canopy overlap
mean(vtm_total_canopy_plot_conifer$percent_canopy_cover)
# 79.10628

canopy_cover_overlap_conifer <- vtm_total_canopy_plot_conifer %>% 
  mutate (national_forest_name = "All Plots") %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

canopy_cover_overlap_forests_conifer <- vtm_total_canopy_plot_conifer %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

canopy_cover_overlap_summary_conifer <- do.call("rbind", list(canopy_cover_overlap_conifer,
                                          canopy_cover_overlap_forests_conifer))

## -----------------------------------------------------------------------------------------------
# without canopy overlap
# average percent canopy cover across all YPMC plots when removing for canopy overlap
mean(vtm_total_canopy_plot_conifer$percent_canopy_cover_nonoverlap)
# 47.26137

canopy_cover_nonoverlap_conifer <- vtm_total_canopy_plot_conifer %>% 
  mutate (national_forest_name = "All Plots") %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

canopy_cover_nonoverlap_forests_conifer <- vtm_total_canopy_plot_conifer %>% 
  group_by(national_forest_name) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

canopy_cover_nonoverlap_summary_conifer <- do.call("rbind", list(canopy_cover_nonoverlap_conifer,
                                          canopy_cover_nonoverlap_forests_conifer))
```


# Make graphs
```{r}
# create final dataframe for graph
canopy_cover_nonoverlap_summary_rm_NA_conifer <-canopy_cover_nonoverlap_summary_conifer %>% 
  mutate(color = case_when(national_forest_name == "All Plots" ~ "1",
                           national_forest_name == "anf" ~ "0",
                           national_forest_name == "cnf" ~ "0",
                           national_forest_name == "lpnf" ~ "0",
                           national_forest_name == "sbnf" ~ "0",
                           national_forest_name == "N/A" ~ "0",
                           TRUE ~ "none"))
canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name == "anf"] <-"ANF"
canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name == "cnf"] <-"CNF"
canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name == "lpnf"] <-"LPNF"
canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name == "sbnf"] <-"SBNF"
canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name == "N/A"] <-"Other"
canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name[canopy_cover_nonoverlap_summary_rm_NA_conifer$national_forest_name == "All Plots"] <-"All YPMC"

# make factors
x_factor <- c("ANF", "CNF", "LPNF", "SBNF","Other", "All YPMC")
color_factor <- c("0", "1")

vtm_grraph_no_oaks <-ggplot(canopy_cover_nonoverlap_summary_rm_NA_conifer,aes(x=factor(national_forest_name, x_factor),y=mean, fill=factor(color, color_factor)))+
   scale_fill_manual(values = c("#636161", "#A00D35"))+
  geom_col(stat="identity",position=position_dodge(width = 0.4), width = 0.4)+
  geom_point(data = canopy_cover_nonoverlap_summary_rm_NA_conifer, aes(x=national_forest_name, y= mean),position=position_dodge(width = 0.4), color = "black", size = 1.3, shape = 16)+
geom_errorbar(data = canopy_cover_nonoverlap_summary_rm_NA_conifer,
              position=position_dodge(width = 0.4),
                aes(x = national_forest_name, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC (Without Oaks)", x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0, 70))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.position = "none")
```

## FIA (without oaks)

# Combine all dataframes to calculate relative canopy cover
```{r}
# combine all species dataframes back together so we can sum the species in plots
fia_total_canopy_conifer <- do.call("rbind", list(fia_abco_sum,
                                          fia_cade_sum,
                                          fia_pico_sum,
                                          fia_pije_sum,
                                          fia_pila_sum,
                                          fia_pipo_sum,
                                          fia_psma_sum))

## -------------------------------------------------------------------------------------------------
# Next we can sum the total canopy cover of all trees from all size classes PER PLOT
fia_total_canopy_plot_conifer <- fia_total_canopy_conifer %>% 
  group_by(unique_plot, admin_forest) %>% 
  tally(n) %>% 
  rename("total_canopy_area" = "n" )

## ------------------------------------------------------------------------------------------------
# right now the total_canopy_area column is in ft^2, we want to convert this to m^2. 
# 1 ft^2 = 0.092903 m^2

# create a column that converts squared feet to squared meters
fia_total_canopy_plot_conifer <- fia_total_canopy_plot_conifer %>% 
  mutate(total_canopy_area_meters = (total_canopy_area *0.092903))

# Divide the total canopy area per plot (measured in m^2) by the size of the plot (also measured in m^2). Note: A vtm plot is 672.45 m^2. Then multiply by 100 to obtain a percent canopy cover. 
fia_total_canopy_plot_conifer <- fia_total_canopy_plot_conifer %>% 
  mutate(percent_canopy_cover = ((total_canopy_area_meters/672.45) * 100))

# note that some of the percent covers are above 100%, because we have not yet accounted for overlapping canopies. 
# we have now determined the percent canopy cover without accounting for overlap--> equivalent to equation 1 in Crookston and Stage 1999 (https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf).
```

# Now apply equation 2 from Crookston and Stage 1999 to account for overlapping canopies. 
```{r}
# Equation 2 (page 2 of https://www.fs.fed.us/rm/pubs/rmrs_gtr024.pdf):
# C = 100 [1 – exp ( – .01 C′ )]
# C = percent canopy cover that accounts for overlap,
# C′ = equation 1-percent canopy cover without accounting for overlap. *This has been calculated in the section above. 
# exp = e^

fia_total_canopy_plot_conifer <- fia_total_canopy_plot_conifer %>% 
  mutate(percent_canopy_cover_nonoverlap = (100*(1-(e^(-.01*percent_canopy_cover)))))
# I THINK it's okay to do this step here (after I've summed the canopy cover per plot), but I can go back and check if applying this equation earlier on changes our results. 
```

# Summarize Findings
```{r}
# with canopy overlap
# average percent canopy cover across all YPMC plots when allowing for canopy overlap
mean(fia_total_canopy_plot_conifer$percent_canopy_cover)
# 116.2319

fia_canopy_cover_overlap_conifer <- fia_total_canopy_plot_conifer %>% 
  mutate (admin_forest = "All Plots") %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

fia_canopy_cover_overlap_forests_conifer <- fia_total_canopy_plot_conifer %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover),
    sd = sd(percent_canopy_cover),
    sample_size = n(),
    se = sd(percent_canopy_cover)/sqrt(n()),
    var = var(percent_canopy_cover))

fia_canopy_cover_overlap_summary_conifer <- do.call("rbind", list(fia_canopy_cover_overlap_conifer,fia_canopy_cover_overlap_forests_conifer))

## -----------------------------------------------------------------------------------------------
# without canopy overlap
# average percent canopy cover across all YPMC plots when removing for canopy overlap
mean(fia_total_canopy_plot_conifer$percent_canopy_cover_nonoverlap)
# 57.42095

fia_canopy_cover_nonoverlap_conifer <- fia_total_canopy_plot_conifer  %>% 
  mutate (admin_forest = "All Plots") %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

fia_canopy_cover_nonoverlap_forests_conifer  <- fia_total_canopy_plot_conifer  %>% 
  group_by(admin_forest) %>% 
  summarize(
    mean = mean(percent_canopy_cover_nonoverlap),
    sd = sd(percent_canopy_cover_nonoverlap),
    sample_size = n(),
    se = sd(percent_canopy_cover_nonoverlap)/sqrt(n()),
    var = var(percent_canopy_cover_nonoverlap))

fia_canopy_cover_nonoverlap_summary_conifer  <- do.call("rbind", list(fia_canopy_cover_nonoverlap_conifer,fia_canopy_cover_nonoverlap_forests_conifer ))
```

# Make Graph 
```{r}
fia_canopy_cover_nonoverlap_summary_rm_NA_conifer  <-fia_canopy_cover_nonoverlap_summary_conifer  %>% 
  mutate(color = case_when(admin_forest == "All Plots" ~ "1",
                           admin_forest == "ANF" ~ "0",
                           admin_forest == "CNF" ~ "0",
                           admin_forest == "LPNF" ~ "0",
                           admin_forest == "SBNF" ~ "0",
                           admin_forest == "Other" ~ "0",
                           TRUE ~ "none"))

# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "anf"] <-"ANF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "cnf"] <-"CNF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "lpnf"] <-"LPNF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "sbnf"] <-"SBNF"
# fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA$admin_forest == "N/A"] <-"Other"
fia_canopy_cover_nonoverlap_summary_rm_NA_conifer $admin_forest[fia_canopy_cover_nonoverlap_summary_rm_NA_conifer $admin_forest == "All Plots"] <-"All YPMC"

# make factors
x_factor <- c("ANF", "CNF", "LPNF", "SBNF","Other", "All YPMC")
color_factor <- c("0", "1")

fia_graph_no_oaks <-ggplot(fia_canopy_cover_nonoverlap_summary_rm_NA_conifer ,aes(x=factor(admin_forest, x_factor),y=mean, fill=factor(color, color_factor)))+
   scale_fill_manual(values = c("#636161", "#A00D35"))+
  geom_col(stat="identity",position=position_dodge(width = 0.4), width = 0.4)+
  geom_point(data = fia_canopy_cover_nonoverlap_summary_rm_NA_conifer , aes(x=admin_forest, y= mean),position=position_dodge(width = 0.4), color = "black", size = 1.3, shape = 16)+
geom_errorbar(data = fia_canopy_cover_nonoverlap_summary_rm_NA_conifer ,
              position=position_dodge(width = 0.4),
                aes(x = admin_forest, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC (Without Oaks)", x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0, 75))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.position = "none") 
```

## Compare (VTM v. FIA without oaks)

```{r}
#combine datasets
#canopy_cover_nonoverlap_summary_rm_NA_conifer vtm data
#canopy_cover_nonoverlap_summary_rm_NA_conifer fia data
canopy_cover_nonoverlap_summary_rm_NA_2_conifer <- canopy_cover_nonoverlap_summary_rm_NA_conifer %>% 
  mutate(plot = case_when(national_forest_name == "ANF" ~ "VTM",
                          national_forest_name == "CNF" ~ "VTM",
                          national_forest_name == "SBNF" ~ "VTM",
                          national_forest_name == "LPNF" ~ "VTM",
                          national_forest_name == "Other" ~ "VTM",
                          national_forest_name == "All YPMC" ~ "VTM_1",
                          TRUE ~ "none"))
  
  
fia_canopy_cover_nonoverlap_summary_rm_NA_2_conifer <- fia_canopy_cover_nonoverlap_summary_rm_NA_conifer %>% 
  mutate(plot = case_when(admin_forest == "ANF" ~ "FIA",
                          admin_forest == "CNF" ~ "FIA",
                          admin_forest == "SBNF" ~ "FIA",
                          admin_forest == "LPNF" ~ "FIA",
                          admin_forest == "Other" ~ "FIA",
                          admin_forest == "All YPMC" ~ "FIA_1",
                          TRUE ~ "none")) %>% 
  rename( "national_forest_name" = "admin_forest")


canopy_cover_comparison_conifer <- do.call("rbind", list(canopy_cover_nonoverlap_summary_rm_NA_2_conifer,
                                          fia_canopy_cover_nonoverlap_summary_rm_NA_2_conifer))

canopy_cover_comparison_conifer <- canopy_cover_comparison_conifer %>%
  filter(national_forest_name %in% c("ANF", "SBNF", "LPNF", "Other", "All YPMC"))

canopy_cover_comparison_2_conifer <- canopy_cover_comparison_conifer %>% 
  filter(national_forest_name == "All YPMC")
```

# make comparison graph
```{r}
x_factor <- c("ANF", "LPNF", "SBNF","Other", "All YPMC")
color_factor <- c("0", "1")
order_factor <- c("VTM", "FIA", "VTM_1", "FIA_1")

ggplot(canopy_cover_comparison_conifer,aes(x=factor(national_forest_name, x_factor),y=mean, fill=factor(plot, order_factor)))+
   scale_fill_manual(values = c("#9C9A9A","#636161", "#BD606E", "#A00D35"), labels=c("VTM", "FIA", "VTM", "FIA"))+
  geom_col(stat="identity",position=position_dodge(width = 0.7), width = 0.7)+
    geom_text(data = canopy_cover_comparison_2_conifer, aes(label=round(mean, 1)), position=position_dodge(width = 0.7), vjust=-1.6, size = 4)+
  geom_point(data = canopy_cover_comparison_conifer, aes(x=national_forest_name, y= mean),position=position_dodge(width = 0.7), color = "black", size = 1.3, shape = 16, show.legend = FALSE)+
geom_errorbar(data = canopy_cover_comparison_conifer,
              position=position_dodge(width = 0.7),
                aes(x = national_forest_name, 
                    ymin = mean - se,
                    ymax = mean + se),
                color = "black",
                width = 0.08) +
  theme_bw()+
  labs(title = "Average Percent Canopy Cover of YPMC (Excluding Oaks)",
       subtitle = "Using mean + 1SD as the midpoint for the 36+ size class",
       x = "Location", y = "Percent Canopy Cover") +
  theme(plot.title = element_text(hjust= 0.5,
                                  face = "bold",
                                  margin=margin(0,0,10,0))) +
  theme(plot.subtitle = element_text(hjust= 0.5,
                                  margin=margin(0,0,10,0))) +
  theme(axis.title.x = element_text (face = "bold",
                                     margin=margin(10,0,0,0))) +
  theme(axis.title.y = element_text (face = "bold",
                                     margin=margin(0,10,0,0)))+
  scale_y_continuous( expand = c(0,0))+
  expand_limits(y = c(0,80))+
  theme(axis.text.x = element_text(color = "black", size = 9))+
  theme(legend.title = element_blank()) +
  ggsave(here::here("NRV Analysis", "canopy_cover", "Figures", "vtm_fia_canopycover_mean+1sd_wo_oaks.png"),
         height = 5,
         width = 7,
         units="in")

```

# percent change of canopy cover (without oaks)
```{r}
pct_change_binned_fia_canopy_wo_oaks <- canopy_cover_comparison_conifer %>%
  select(national_forest_name, mean, plot) %>%
  group_by(national_forest_name) %>%
  mutate(pct_change = ((mean/lag(mean) - 1)*100))

# pct_change_binned_fia <- canopy_cover_comparison %>%
#   select(national_forest_name, mean, plot) %>%
#   group_by(national_forest_name) %>%
#   mutate(pct_change = ((mean/lag(mean) - 1)*100)) 
```

