---
title: "USFS FIA Tree Data Review"
author: "Jennifer Truong"
date: "9/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

# Attach packages ----

library(tidyverse)
library(janitor)
library(naniar)
library(kableExtra)
library(lubridate)
library(here)

```

```{r}
# Read in data ----

# Read in CSVs from all four national forests

anf_fia <- read_csv("G:/Github/gotforestry/R/fia_data_observation/tree_data/ANF_FSVEG_TXPT_V.csv") %>% 
  clean_names() # Change to lower case

cnf_fia <- read_csv("G:/Github/gotforestry/R/fia_data_observation/tree_data/CNF_FSVEG_TXPT_V.csv") %>% 
  clean_names() # Change to lower case

lpf_fia <- read_csv("G:/Github/gotforestry/R/fia_data_observation/tree_data/LPF_FSVEG_TXPT_V.csv") %>% 
  clean_names() %>% # Change to lower case
  mutate(
    plot = as.character(plot) # Change plot column from double to character because it's numeric for some reason
  )

bdf_fia <- read_csv("G:/Github/gotforestry/R/fia_data_observation/tree_data/BDF_FSVEG_TXPT_V.csv") %>% 
  clean_names() %>% # Change to lower case
  mutate(
    forest = as.character(forest) # Change forest column from double to character because it's numeric for some reason
  )

# Combine all CSVs together
all_nfs_fia <- bind_rows(anf_fia, cnf_fia, lpf_fia, bdf_fia) %>% 
  rename("species" = "spec") %>% # Rename species column
  separate(
    col = ddate,
    into = c("ddate_month", "ddate_year"), # Create two new columns
    sep = -4, # Split 4 digits from the right
    convert = TRUE
  ) %>% 
  mutate(
    ddate_year = as.character(ddate_year) # Change year column to character 
  ) %>% 
  mutate(
    species = case_when(
      species %in% "CADE27" ~ "CADE", # Change FSVEG code for Calocedrus decurrens to CADE
      species %in% c("PICO", "PICOM") ~ "PICOM", # Change FSVEG codes for Pinus contorta to PICOM <- important to do this first before the next line
      species %in% "PICO3" ~ "PICO", # Change FSVEG code for Pinus coulteri to PICO
      species %in% c("QUCH2", "QUCHC") ~ "QUCH", # Change FSVEG code for Quercus chrysolepis to QUCH
      TRUE ~ as.character(species) # Keep all other species the same
      )
  )


# FIA Metadata notes ----

# Only limited to National Forest data, no FIA plots outside of NF boundaries included in this dataset

# Additional notes ----

# To clean ddate column as actual dates, split column into two and have last four digits as year by using separate()
# Troubleshooting help: https://tidyr.tidyverse.org/reference/separate.html


```

```{r Quick analysis of CADE in LPF}

lpf_cade <- lpf_fia %>% 
  filter(spec == "CADE27") %>% #Filter to CADE
  group_by(plot) %>% 
  count()
```

# Preliminary Data Analysis

### Observing Unique Species

YPMC Conifer species we are focusing on for our GP:

| Scientific Name               | Common Name          | Code       |
| ----------------------------- | -------------------- | ---------- |
| Pinus ponderosa               | Ponderosa pine       | PIPO       |
| Pinus jeffreyi                | Jeffery pine         | PIJE       |
| Pinus lambertiana             | Sugar pine           | PILA       |
| Abies concolor                | White fir            | ABCO       |
| Calocedrus decurrens          | Incense cedar        | CADE       |

Affiliate species that we may look at:

| Scientific Name               | Common Name          | Code       |
| ----------------------------- | -------------------- | ---------- |
| Pinus coulteri                | Coulter pine         | PICO3      |
| Pseudotsuga macrocarpa        | Big cone douglas fir | PSMA       |
| Quercus kelloggii             | California black oak | QUKE       |
| Quercus chrysolepis           | Canyon live oak      | QUCH       |

Conifer species we are not considering for our analysis:

| Scientific Name               | Common Name          | Code       |
| ----------------------------- | -------------------- | ---------- |
| Pinus contorta ssp. murrayana | Lodgepole pine       | PICOM      | 

### Setup for Filtering Plots by Species

```{r Species Filtering Setup}

# Set up vector for YPMC species
alpine = c("PIPO", "PIJE", "PILA", "ABCO", "CADE")

# Set up vector for "PSMA", "PICO3", "QUKE", "QUCH", and "PICOM"
subalpine = c("PSMA", "PICO", "QUKE", "QUCH", "PICOM")


```

### Tree Size Classes

Diameter base height in cm; classes determined from Figure 19 (pg 106 hard-copy and 116 digital) in Safford and Stevens 2017

Use variable DBHI and convert from inches to cm

- 5 - 20 cm 
- 20 - 40 cm
- 40 - 60 cm
- 60 - 80 cm
- 80 - 100 cm
- 100 - 120 cm
- 120 - 140 cm
- /> 140 cm

# FIA Data Analysis by NF

### Data Collection Dates

```{r Draw timeline of data connection}

# Create new df for graph
all_nfs_fia_timeline_df <- all_nfs_fia %>% 
  group_by(forest, ddate_year) %>% # Group by forest and year
  count() # Count # of rows

# Try to draw a timeline (?)
ggplot(data = all_nfs_fia_timeline_df, 
       aes(x = ddate_year,
           y = n)) +
  geom_line(aes(color = forest, # Color by forest
                group = forest) # Group observations by forest
            ) + # Change line thickness to 1
  scale_color_manual(name = "National Forest",
                     labels = c("Angeles", "Cleveland", "Los Padres", "San Bernardino"), # Rename legend items to forest names
                     values = c("coral", "darkgoldenrod2", "darkgreen", "deepskyblue2")) + # Specify colors
  labs(x = "Year",
       y = "Number of Observations",
       title = "FIA Observations per Year") + # Change axis labels and title
  theme_light() +
  theme(
    plot.title = element_text(hjust = 0.5) # Center plot title
  )

# Note: FIA record dates vary from 2001 - 2011 for all Nfs
```

Los Padres seems to have the most observations overall. The highest amount of observations occurred in 2011 with 6000 observations total. This wsas followed by San Bernardino with about 4000 observations in 2010.

### Number of FIA Observations

```{r Count of Observations}

# Get total number of observations for all forests
all_nfs_fia_total <- nrow(all_nfs_fia)

# Get number of observations by forest
fia_total_per_forest <- all_nfs_fia %>% 
  group_by(forest) %>% # Group by forest
  count() %>% # Count number of rows
  mutate(
    forest = case_when(
      forest == "01" ~ "Angeles",
      forest == "02" ~ "Cleveland",
      forest == "03" ~ "Los Padres",
      forest == "04" ~ "San Bernardino"
    )
  ) # Change values in forest column to forest names

# Create table
fia_total_per_forest %>%
  kable() %>% 
  kable_styling(full_width = F)

```

### YPMC Conifer Count

```{r Count observations per species}

# Find all tree species recorded in "species" column ----
fia_all_species <- unique(all_nfs_fia$species)
# Troubleshooting tip for unique(): https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/

# Count number of observations by species per forest
fia_species_per_forest <- all_nfs_fia %>% 
  group_by(forest, species) %>%   # Group by forest and species
  count() # Count observations per species

# Filter to species of interest
fia_species_count <- fia_species_per_forest %>% 
  filter(species %in% c(alpine, subalpine))
# Troubleshooting tip for using %in% for a groups of values: https://www.datanovia.com/en/lessons/subset-data-frame-rows-in-r/

# Get sum for conifer observations
fia_species_sum <- sum(fia_species_count$n)
# Troubleshooting tip for summing values of one variable: https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/sum

# Make table for conifer count ----
fia_species_count %>%
  kable() %>% 
  kable_styling(full_width = F)

```

### Number of Trees per Plot

```{r Grouping Data Points by Plot}

# First, group the data by plot number and count
fia_plot_count <- all_nfs_fia %>% 
  group_by(forest, plot) %>% 
  count()

# Then count number of rows to get number of plots
fia_plot_count_total <- NROW(fia_plot_count)
```

There are `r fia_plot_count_total` plots in all four Southern California national forests.

```{r Counting by plot}

# Now count total number of trees per plot (before any filtering)
fia_tree_per_plot <- all_nfs_fia %>% 
  group_by(gis_link) %>% # Group by gis_link. May seem to be more accurate than grouping forest and plot separately
  count()

# Count total number of trees (alpine and subalpine only) per plot
fia_conifer_per_plot <- all_nfs_fia %>% 
  group_by(gis_link) %>% # Group by gis_link (GROUP BY SPECIES IN THE NEXT LINE)
  #separate( 
    #col = gis_link,
    #into = c("gislink_prefix", "r5plotnum"), # Create two new columns
    #sep = -5, # Split 5 digits to recreate the r5plotnum column from the FIA shapefile
    #convert = FALSE
  #) %>% 
  filter(species %in% c(alpine, subalpine)) %>% 
  count()

# Then, do it again by grouping by plot number and by species
fia_plot_species <- all_nfs_fia %>% 
  group_by(gis_link, species) %>% # Group by forest, plot, species
  filter(species %in% c(alpine, subalpine)) %>% 
  count()

# Join the two tables together ----
# Should total # per plot include ALL species or just species of interest? <- ask AMP and Leana
fia_plot_species_join <- inner_join(fia_plot_species, fia_conifer_per_plot, by = "gis_link") %>% 
  rename(
    "species_n" = "n.x",
    "total_n" = "n.y"
  ) %>% 
  mutate(
    pct = round(species_n/total_n*100, 2)
  )

# Summarize tree per species (for all, not just YPMC) for each plot ----
fia_plot_species_join %>% 
  kable() %>% 
  kable_styling(latex_options = "striped",
                full_width = F) # %>% 
  # scroll_box(width = "100px", height = "300px")
# Troubleshooting tip for kableExtra functions: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#html_only_features

```

```{r Just for observation only}
# Filter to LPF CADE for observation purposes

fia_plot_species_join_cade <- fia_plot_species_join %>% 
  filter(gis_link %in% c("LPFPLUS11_107251", "LPFPLUS11_107303", "LPFPLUS11_107402", "LPFPLUS11_107433", "LPFPLUS11_107721", "LPFPLUS11_107722", "LPFPLUS11_107732", "LPFPLUS11_107749"))

fia_plot_species_join_cade %>% 
  kable() %>% 
  kable_styling(latex_options = "striped",
                full_width = F)

# Filter to YPMC plots (for observation purposes, won't appear in final PDF) ----

# Filter to plots with YPMC conifers
fia_plot_conifer_count <- fia_plot_species %>% 
  filter(species %in% alpine) %>%  # Filter to just rows with YPMC species
  group_by(gis_link) %>%  # Group by forest and plot
  mutate(
    conifer_total = sum(n) # Sum observations of conifers per plot
  )

# Filter joined table to just alpine
fia_plot_join_conifer <- fia_plot_species_join %>% 
  filter(species %in% alpine)

```

#### Notes about Plots

##### Angeles

1. Plot 01020 has 1 CADE and 1 QUCH according to the table. Matches what's in the final DF too <- Should I remove it?
2. Plot 01024 has 15 QUCH according to the table. However, in the final DF, more rows are present and they seem to be all YPMC species. I wonder what happened. **Edit:** Grouping by "gis_link" seems to get rid of the disrepancy. Plot 01024 has been removed from the final DF. 

### Tree Size Class for all NFs

```{r Categorizing by Tree Size Class}

fia_size_class <- all_nfs_fia %>% 
  filter(species %in% alpine) %>%  # Filter by conifer species
  mutate(
    dbh_cm = dbhi*2.54 # Convert inches to cm
  )

#Create new vector for size classes

fia_size_class$size_class <- cut(fia_size_class$dbh_cm,
                                 c(0,5,20,40,60,80,100,120,140,200),
                                 labels = c("0-5", "5-20", "20-40", "40-60", "60-80", "80-100", "100-120", "120-140", ">140"))

# Count per size class

fia_size_class_count <- fia_size_class %>% 
  group_by(forest, size_class) %>% 
  count(size_class) 

fia_size_class_count %>%    
  kable(col.names = c("Forest", "Size Class (cm)", "Count")) %>% 
  kable_styling(latex_options = "striped",
                full_width = F) # %>% 
  #scroll_box(width = "300px", height = "300px")

# Count per size class and conifer species

fia_size_class_conifer_count <- fia_size_class %>% 
  group_by(forest, size_class, species) %>% 
  count(size_class)

fia_size_class_conifer_count %>% 
  kable(col.names = c("Forest", "Size Class (cm)", "Species", "Count")) %>% 
  kable_styling(latex_options = "striped",
                full_width = F) #%>% 
  #scroll_box(width = "300px", height = "300px")

# Troubleshooting Tip - creating catagories based on ranges: https://stackoverflow.com/questions/21050021/create-category-based-on-range-in-r
```
 


### Finalizing of FIA Data with YPMC Data

```{r Write Final CSV for ANF}
# Remove plots from main FIA dataset
all_nfs_fia_ypmc <- all_nfs_fia %>% 
  filter(species %in% c(alpine, subalpine)) %>% # Filter down to alpine and subalpine species
  mutate(
    temp = case_when(
      species %in% c("PSMA", "PICO", "QUKE", "QUCH", "PICOM") ~ "0", 
      TRUE ~ "1") 
  ) %>% 
    mutate(temp = as.numeric(temp)) %>% 
    group_by(gis_link) %>% # Group by gis_link
    mutate(all_plot_species2 = sum(temp)) %>% # Sum percentage of species consumption
    filter(all_plot_species2 != 0) %>%  # Keep  plots that do not equal 0
    select(-temp, -all_plot_species2) # Keep all columns except the temporary columns that were created now

# Write CSV
write_csv(all_nfs_fia_ypmc, "G:/Github/gotforestry/R/fia_final_data/all_nfs_fia_ypmc.csv")

```

#### Plots in final DF:

```{r Pull final plots from final CSV}

all_nfs_fia_ypmc_plots <- all_nfs_fia_ypmc %>% 
  select(forest, plot, gis_link) %>%  # Select forest, plot, and gis_link to verify
  group_by(forest, plot) %>% 
  count()

# Create vector for plots included
all_nfs_fia_plot_vector <- all_nfs_fia_ypmc$gis_link

# Create table for plots included
all_nfs_fia_ypmc_plots %>%    
  kable() %>% 
  kable_styling(latex_options = "striped",
                full_width = F) # %>% 
  #scroll_box(width = "300px", height = "300px")


```


#### Plots removed from final DF: 

```{r Pull removed plots that are not included in final CSV}

# Code not working so far, not sure how much this is important, although it helps with verification

#all_nfs_fia_removed_plots <- all_nfs_fia %>% 
  #filter(species %in% c(alpine, subalpine)) %>% # Filter to alpine and subalpine species
  #select(forest, plot, gis_link) %>%  # Select forest, plot, and gis_link to verify
  #filter(gis_link %in% all_nfs_fia_plot_vector) %>% # Remove plots that match YPMC gis_link values
  #group_by(forest, plot) %>% 
  #count()

# Create table for plots NOT included

#all_nfs_fia_removed_plots %>%    
  #kable() %>% 
  #kable_styling(latex_options = "striped",
                #full_width = F) # %>% 
  #scroll_box(width = "300px", height = "300px")


```
Notes: the notes below are for when I manually looked through the data frames. I didn't finish because it took forever.

- **Angeles:** 01010, 01013, 01014, 01017, 01019, 01024, 01030, 01040, 01043, 01044, 01076, 01077, 01081, 01101, 01104, 01110, 01115, 01116, 01118
- Cleveland:
- Los Padres:
- San Bernardino: