---
title: "FIA Data Review"
author: "Jennifer Truong"
date: "6/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

# Attach packages ----

library(tidyverse)
library(janitor)
library(naniar)
library(kableExtra)
library(lubridate)

```

```{r}
# Read in data ----

# Cleveland NF
cnf_fia <- read_csv(here::here ("R", "fia_data_observation", "plot_data", "CNF_FSVEG_VXPT_V.csv")) %>% 
  clean_names() %>%  # Clean names
  rename("species" = "spec") %>% # Rename species column
  separate(
    col = ddate,
    into = c("ddate_month", "ddate_year"), # Create two new columns
    sep = -4, # Split 4 digits from the right
    convert = TRUE
)

# Angeles NF
anf_fia <- read_csv(here::here ("R", "fia_data_observation", "plot_data", "ANF_FSVEG_VXPT_V.csv")) %>% 
  clean_names() %>% 
  rename("species" = "spec") %>% 
  separate(
    col = ddate,
    into = c("ddate_month", "ddate_year"), # Create two new columns
    sep = -4, # Split 4 digits from the right
    convert = TRUE
  )

# Los Padres NF
lpf_fia <- read_csv(here::here ("R", "fia_data_observation", "plot_data","LPF_FSVEG_VXPT_V.csv")) %>% 
  clean_names() %>% 
  rename("species" = "spec") %>% 
  separate(
    col = ddate,
    into = c("ddate_month", "ddate_year"), # Create two new columns
    sep = -4, # Split 4 digits from the right
    convert = TRUE
  )

# San Bernardino NF
bdf_fia <- read_csv(here::here ("R", "fia_data_observation", "plot_data","BDF_FSVEG_VXPT_V.csv")) %>% 
  clean_names() %>% 
  rename("species" = "spec") %>% 
  separate(
    col = ddate,
    into = c("ddate_month", "ddate_year"), # Create two new columns
    sep = -4, # Split 4 digits from the right
    convert = TRUE
  )
# FIA Metadata notes ----

# Link: https://www.fs.fed.us/r5/rsl/projects/inventory/fsveg-veg.shtml

# According to FIA metadata, the following conditions apply to conifer species
# - Special Features (SPFE) and Surface Cover (SRFC) are not applicable so they'll be NA anyways
# - Number of Stems (NCNT) is optional

# SRFC refers to non-vegetative cover (ie. rocks, water/snow/ice, moss/lichen/fungi, etc).
# NCNT is just number of cover stems...still have no idea what that means.

# Note: Every time there is an NA value in "spec," there is a value in "srfc" and vice versa. Tree/plant =/= surface cover?

# Additional notes ----

# To clean ddate column as actual dates, split column into two and have last four digits as year by using separate()
# Troubleshooting help: https://tidyr.tidyverse.org/reference/separate.html

```

## Observing Unique Species

Conifer species we are focusing on for our GP:

| Scientific Name               | Common Name          | Code       |
| ----------------------------- | -------------------- | ---------- |
| Pinus ponderosa               | Ponderosa pine       | PIPO       |
| Pinus jeffreyi                | Jeffery pine         | PIJE       |
| Pinus coulteri                | Coulter pine         | PICO3      |
| Pinus lambertiana             | Sugar pine           | PILA       |
| Pseudotsuga macrocarpa        | Big cone douglas fir | PSMA       |
| Abies concolor                | White fir            | ABCO       |
| Calocedrus decurrens          | Incense cedar        | CADE       |
| Pinus contorta ssp. murrayana | Lodgepole pine       | PICO/PICOM | 


## ANF Data Analysis

### ANF Data Collection Dates

```{r Draw timeline of data connection}

# anf_date <- as_date(anf_fia$ddate); was unable to parse column as dates

# Try to draw a timeline (?)
ggplot(data = anf_fia, aes(x = ddate_year)) +
  geom_density()
```

### ANF Conifer Count

```{r ANF Species}

# Get total number of observations for ANF ----

anf_total <- nrow(anf_fia)

# Find all tree species recorded in "species" column ----

anf_all_species <- unique(anf_fia$species)
# Troubleshooting tip for unique(): https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/

anf_all_species_row <- NROW(anf_all_species)
# Troubleshooting tip for counting number of rows: https://stats.stackexchange.com/questions/5253/how-do-i-get-the-number-of-rows-of-a-data-frame-in-r

# Count number of species within ANF ----

anf_count <- anf_fia %>% 
  group_by(species) %>%   # Group by species
  count()       # Count observations per species

# Filter to conifer species
anf_count_conifer <- anf_count %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27"))
# Troubleshooting tip for using %in% for a groups of values: https://www.datanovia.com/en/lessons/subset-data-frame-rows-in-r/

# Get sum for conifer observations
anf_sum_conifer <- sum(anf_count_conifer$n)
# Troubleshooting tip for summing values of one variable: https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/sum

# Make table for conifer count ----
anf_count_conifer %>%
  kable() %>% 
  kable_styling()

```

### ANF Trees per Plot

```{r Grouping Data Points by Plot}

# First, group the data by plot number and count?

anf_plot <- anf_fia %>% 
  group_by(plot) %>% 
  count()

# Then count number of rows to get number of plots

anf_plot_count <- NROW(anf_plot)
```

```{r}
# Then, do it again by grouping by plot number and filtering to species

anf_plot_conifer <- anf_fia %>% 
  group_by(plot) %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27")) %>% 
  count()

anf_plot_conifer %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "100px", height = "300px")
# Troubleshooting tip for kableExtra functions: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#html_only_features

# Doing it this way excludes plots that have other tree species or no trees recorded at all.

```

### ANF Missing Data

```{r Counting NA Values}
# Observing NA values for ANF ----

# Total Count of NA values - 80259
anf_na <- n_miss(anf_fia)

# Total Count of complete values - 194861
anf_complete <- n_complete(anf_fia)

# Percentage of NA values - 29.17%
anf_na_pct <- pct_miss(anf_fia)

# Summary of NA values
anf_na_summ <- miss_var_summary(anf_fia)

# Turn summary table into kableExtra table
anf_na_summ %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "300px", height = "300px")

# Note #1: NCNT has 0 NA values here...does that mean those 0's represent 0's or are some of them NA values?
# Note #2: This summary would be more or less the same for the other forests.

# Summary of NA values by observations (rows)
anf_na_summ_row <-  miss_case_table(anf_fia)
```

```{r ANF Other Tree Data}
# Find difference between total observations and conifer trees (excluding NA values in species column)

anf_other_trees <- (anf_total - anf_na_summ$n_miss[14]) - anf_sum_conifer

```

### ANF FIA Data Summary 

- The total number of data points recorded in the FIA data for ANF is `r anf_total`. There are `r anf_na_summ$n_miss[11]` data points recorded as actual plant trees.
- There are `r anf_plot_count` plots in ANF.
- A total of `r anf_all_species_row` unique plant species are recorded in this FIA dataset. Only 8 of these are our conifer species of interest. 
- `r anf_sum_conifer` trees are recorded as conifer trees. 
- `r anf_other_trees` are recorded as other plant species.
- There are `r anf_na_summ$n_miss[14]` missing values in the species column.

## CNF Data Analysis

### CNF Data Collection Dates

```{r}

ggplot(data = cnf_fia, aes(x = ddate_year)) +
  geom_density()
```

### CNF Conifer Count
```{r CNF Species}

# Get total number of observations for CNF ----

cnf_total <- nrow(cnf_fia)

# Find all tree species recorded in "species" column ----

cnf_all_species <- unique(cnf_fia$species)
# Troubleshooting tip for unique(): https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/

cnf_all_species_row <- NROW(cnf_all_species)
# Troubleshooting tip for counting number of rows: https://stats.stackexchange.com/questions/5253/how-do-i-get-the-number-of-rows-of-a-data-frame-in-r

# Count number of species within CNF ----

cnf_count <- cnf_fia %>% 
  group_by(species) %>%   # Group by species
  count()       # Count observations per species

# Filter to conifer species
cnf_count_conifer <- cnf_count %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27"))
# Troubleshooting tip for using %in% for a groups of values: https://www.datanovia.com/en/lessons/subset-data-frame-rows-in-r/

# Get sum for conifer observations
cnf_sum_conifer <- sum(cnf_count_conifer$n)
# Troubleshooting tip for summing values of one variable: https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/sum

# Make table for conifer count ----
cnf_count_conifer %>%
  kable() %>% 
  kable_styling()

```

### CNF Trees per Plot

```{r}

# First, group the data by plot number and count?

cnf_plot <- cnf_fia %>% 
  group_by(plot) %>% 
  count()

# Then count number of rows

cnf_plot_count <- NROW(cnf_plot)
```

```{r}
# Then, do it again by grouping by plot number and filtering to species

cnf_plot_conifer <- cnf_fia %>% 
  group_by(plot) %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27")) %>% 
  count()

cnf_plot_conifer %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "100px", height = "300px")
# Troubleshooting tip for kableExtra functions: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#html_only_features

# Doing it this way excludes plots that have other tree species or no trees recorded at all.

```

### CNF Missing Data

```{r}
# Observing NA values for CNF ----

# Total Count of NA values - 55371
cnf_na <- n_miss(cnf_fia)

# Total Count of complete values - 131989
cnf_complete <- n_complete(cnf_fia)

# Percentage of NA values - 29.55%
cnf_na_pct <- pct_miss(cnf_fia)

# Summary of NA values
cnf_na_summ <- miss_var_summary(cnf_fia)

cnf_na_summ %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "300px", height = "300px")

# Summary of NA values by observations (rows)
cnf_na_summ_row <-  miss_case_table(cnf_fia)
```

```{r CNF Other Tree Data}
# Find difference between total observations and conifer trees (excluding NA values in species column)

cnf_other_trees <- (cnf_total - cnf_na_summ$n_miss[14]) - cnf_sum_conifer

```

### CNF FIA Data Summary 

- The total number of data points recorded in the FIA data for CNF is `r cnf_total`. There are `r cnf_na_summ$n_miss[11]` data points recorded as actual plant trees.
- There are `r cnf_plot_count` plots in CNF.
- A total of `r cnf_all_species_row` unique plant species are recorded in this FIA dataset. Only 8 of these are our conifer species of interest. 
- `r cnf_sum_conifer` trees are recorded as conifer trees. 
- `r cnf_other_trees` are recorded as other plant species.
- There are `r cnf_na_summ$n_miss[14]` missing values in the species column.

## LPF Data Analysis

### LPF Data Collection Dates

```{r}

ggplot(data = lpf_fia, aes(x = ddate_year)) +
  geom_density()
```

### LPF Conifer Count

```{r}

# Get total number of observations for LPF ----

lpf_total <- nrow(lpf_fia)

# Find all tree species recorded in "species" column ----

lpf_all_species <- unique(lpf_fia$species)
# Troubleshooting tip for unique(): https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/

lpf_all_species_row <- NROW(lpf_all_species)
# Troubleshooting tip for counting number of rows: https://stats.stackexchange.com/questions/5253/how-do-i-get-the-number-of-rows-of-a-data-frame-in-r

# Count number of species within LPF ----

lpf_count <- lpf_fia %>% 
  group_by(species) %>%   # Group by species
  count()       # Count observations per species

# Filter to conifer species
lpf_count_conifer <- lpf_count %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27"))

# Get sum for conifer observations
lpf_sum_conifer <- sum(lpf_count_conifer$n)

# Make table for conifer count
lpf_count_conifer %>%
  kable() %>% 
  kable_styling()
```

### LPF Trees per Plot

```{r}

# First, group the data by plot number and count?

lpf_plot <- lpf_fia %>% 
  group_by(plot) %>% 
  count()

# Then count number of rows

lpf_plot_count <- NROW(lpf_plot)
```

```{r}
# Then, do it again by grouping by plot number and filtering to species

lpf_plot_conifer <- lpf_fia %>% 
  group_by(plot) %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27")) %>% 
  count()

lpf_plot_conifer %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "100px", height = "300px")
# Troubleshooting tip for kableExtra functions: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#html_only_features

# Doing it this way excludes plots that have other tree species or no trees recorded at all.

```

### LPF Missing Data

```{r}
# Observing NA values for LPF ----

# Total Count of NA values - 246437
lpf_na <- n_miss(lpf_fia)

# Total Count of complete values - 565323
lpf_complete <- n_complete(lpf_fia)

# Percentage of NA values - 30.36%
lpf_na_pct <- pct_miss(lpf_fia)

# Summary of NA values
lpf_na_summ <- miss_var_summary(lpf_fia)

lpf_na_summ %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "300px", height = "300px")

# Summary of NA values by observations (rows)
lpf_na_summ_row <-  miss_case_table(lpf_fia)
```

```{r LPF Other Tree Data}
# Find difference between total observations and conifer trees (excluding NA values in species column)

lpf_other_trees <- (lpf_total - lpf_na_summ$n_miss[13]) - lpf_sum_conifer

```

### LPF FIA Data Summary

- The total number of data points recorded in the FIA data for LPF is `r lpf_total`. There are `r lpf_na_summ$n_miss[14]` data points recorded as actual plant trees.
- There are `r lpf_plot_count` plots in LPF.
- A total of `r lpf_all_species_row` unique plant species are recorded in this FIA dataset. Only 8 of these are our conifer species of interest. 
- `r lpf_sum_conifer` trees are recorded as conifer trees. 
- `r lpf_other_trees` are recorded as other plant species.
- There are `r lpf_na_summ$n_miss[13]` missing values in the species column.

## BDF Data Analysis

```{r}

ggplot(data = bdf_fia, aes(x = ddate_year)) +
  geom_density()
```

### BDF Conifer Count

```{r}

# Get total number of observations for BDF ----

bdf_total <- nrow(bdf_fia)

# Find all tree species recorded in "species" column ----

bdf_all_species <- unique(bdf_fia$species)
# Troubleshooting tip for unique(): https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/

bdf_all_species_row <- NROW(bdf_all_species)
# Troubleshooting tip for counting number of rows: https://stats.stackexchange.com/questions/5253/how-do-i-get-the-number-of-rows-of-a-data-frame-in-r

# Count number of species within BDF ----

bdf_count <- bdf_fia %>% 
  group_by(species) %>%   # Group by species
  count()       # Count observations per species

# Filter to conifer species
bdf_count_conifer <- bdf_count %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27"))

# Get sum for conifer observations
bdf_sum_conifer <- sum(bdf_count_conifer$n)

# Make table for conifer count
bdf_count_conifer %>%
  kable() %>% 
  kable_styling()
```

### BDF Trees per Plot

```{r}

# First, group the data by plot number and count?

bdf_plot <- bdf_fia %>% 
  group_by(plot) %>% 
  count()

# Then count number of rows

bdf_plot_count <- NROW(bdf_plot)
```

```{r}

# Then, do it again by grouping by plot number and filtering to species

bdf_plot_conifer <- bdf_fia %>% 
  group_by(plot) %>% 
  filter(species %in% c("PIPO", "PIJE", "PICO", "PICOM", "PICO3", "PILA", "PSMA", "ABCO", "CADE27")) %>% 
  count()

bdf_plot_conifer %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "100px", height = "300px")

# Doing it this way excludes plots that have other tree species or no trees recorded at all.

```

### BDF Missing Data

```{r}
# Observing NA values for BDF ----

# Total Count of NA values - 145939
bdf_na <- n_miss(bdf_fia)

# Total Count of complete values - 345021
bdf_complete <- n_complete(bdf_fia)

# Percentage of NA values - 29.72%
bdf_na_pct <- pct_miss(bdf_fia)

# Summary of NA values
bdf_na_summ <- miss_var_summary(bdf_fia)

bdf_na_summ %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed")) %>% 
  scroll_box(width = "300px", height = "300px")

# Summary of NA values by observations (rows)
bdf_na_summ_row <-  miss_case_table(bdf_fia)

```

```{r BDF Other Tree Data}
# Find difference between total observations and conifer trees (excluding NA values in species column)

bdf_other_trees <- (bdf_total - bdf_na_summ$n_miss[14]) - bdf_sum_conifer

```

### BDF FIA Data Summary

- The total number of data points recorded in the FIA data for BDF is `r bdf_total`. There are `r bdf_na_summ$n_miss[12]` data points recorded as actual plant trees.
- There are `r bdf_plot_count` plots in BDF.
- A total of `r bdf_all_species_row` unique plant species are recorded in this FIA dataset. Only 8 of these are our conifer species of interest. 
- `r bdf_sum_conifer` trees are recorded as conifer trees. 
- `r bdf_other_trees` are recorded as other plant species.
- There are `r bdf_na_summ$n_miss[14]` missing values in the species column.

## Summary of NA Statistics

```{r Summary of NA statistics}
# Combining NA stats together in one df ----

# Create vectors
nf_name <- c("ANF", "CNF", "LPF", "BDF")
nf_na <- c(anf_na, cnf_na, lpf_na, bdf_na)
nf_complete <- c(anf_complete, cnf_complete, lpf_complete, bdf_complete)
nf_na_pct <- c(anf_na_pct, cnf_na_pct, lpf_na_pct, bdf_na_pct)

# Combine vectors into data frame
na_df <- tibble(nf_name, nf_na, nf_complete, nf_na_pct)

# Turn data frame into kable table for a neater look
na_df %>% 
  kable(digits = 2) %>% # Keep deimal points to two places
  kable_styling(bootstrap_options = c("striped", "condensed"))

```

